---
title: "Beyond_CompleteCode"
author: "Brenda Morales"
date: "`r Sys.Date()`"
output: pdf_document
---

## Libraries' importation

```{r}

library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(tibble)
library(ggplot2)
library(MASS)
library(writexl)
library(scales)
library(patchwork)
library(apollo)
library(readxl)
library(psych)
library(car)
library(lavaan)
library(semPlot)
library(gridExtra)
library(semTools)
library(mvnfast)
library(parameters)
library(MVN)
library(plspm)
library(gmnl)

```

## Data importation

```{r}
data_path <- file.path("data", "Analisis.xlsx")

if (!file.exists(data_path)) {
  stop("Data file not found. Place the restricted workbook at data/Analisis.xlsx before running the analysis.")
}

# Measurement model data used in the PLS-SEM sections.
RHC <- read_excel(data_path, sheet = "RHC")
Spain <- read_excel(data_path, sheet = "Spain")
```

# Partial Least Squares - Measurement models
## Registry sample

### Internal consistency
```{r}
psych::alpha((RHC[10:28]))
```

### Confirmatory Factor Analysis

#### KMO and Fligner-Killeen test

```{r}
KMO((RHC[10:27]))
fligner.test((RHC[10:27]))
```
#### Factor loadings

```{r}
scree((RHC[10:26]), main="Sediment graph")
#Factor loadings
Behaviour <- psych::fa(RHC[10:26], nfactor = 3,rotate = "varimax", fm="pa") %>%
model_parameters(sort = TRUE, threshold = "max")

```
### Measurement model

```{r}

path_matrix <- rbind(
  c(0, 0, 0, 0, 0), # Risk
  c(1, 0, 0, 0, 0), # Attitudes
  c(0, 1, 0, 0, 0), # PerceivedBehavC
  c(0, 0, 0, 0, 0), # SubjectiveNorms
  c(1, 1, 1, 1, 0)  # Behaviour
)

#first stage
colnames(path_matrix) <- rownames(path_matrix) <- c("Risk", "Attitudes", "PerceivedBehavC",   "SubjectiveNorms", "Behaviour")

# Measure model
#1st stage
blocks1 <- list(
 c("Rsk1", "Rsk2", "Rsk3", "Rsk4"),  # Risk
  c("Att1", "Att2", "Att3", "Att4"), # Attitudes: Att3 shows high load in PBC so we delete it
  c( "Pbc1", "Pbc2", "Pbc3", "Pbc4"), # PerceivedBehavC: PBC2 shows very low loading in PBC and higher in Risk (delete)
  c("Sn1", "Sn2", "Sn3"),       # SubjectiveNorms
  c("DecIntention", "Offset", "PastRedAction")                        # Behaviour
)

#2nd stage
blocks2 <- list(
  c( "Rsk2", "Rsk3", "Rsk4"), # Risk: Risk1 has low load but not high in any other construct (must check AVE and Fornell Larcker to see if Risk is well adjust)
  c("Att1",  "Att2", "Att4"), # Attitudes: Att3 shows high load in PBC so we delete it
  c("Pbc1", "Pbc3", "Pbc4"),         # PerceivedBehavC # PerceivedBehavC: PBC1 shows very low loading in PBC and highein Risk
  c("Sn1", "Sn2", "Sn3"),       # SubjectiveNorms
  c("DecIntention", "Offset")    # Behaviour: PastRedAction presents a very low load so is deleted
)

#3rd stage
blocks <- list(
  c( "Rsk2", "Rsk3", "Rsk4"), # Risk
  c("Att1",  "Att2", "Att4"),   # Attitudes
  c( "Pbc3", "Pbc4"),    # PerceivedBehavC
  c("Sn1", "Sn2", "Sn3"),       # SubjectiveNorms
  c("DecIntention", "Offset")   # Behaviour
)

# Define all variables as reflexive
modes <- c("A", "A", "A", "A", "A")

#PLS-PM
pls_model <- plspm(RHC, path_matrix, blocks, modes)
summary(pls_model)

crossloadings <- pls_model$crossloadings  %>% dplyr::select( -block)

# Crear tabla LaTeX con xtable
xtable(crossloadings, caption = "Cross loadings PLS", label = "tab:loadings")




crossloadings_named <- crossloadings
rownames(crossloadings) <- crossloadings$name
crossloadings$name <- NULL
crossloadings$block <- NULL  # eliminar si no quieres mostrar 'block'

print(xtable(crossloadings,
             caption = "Cross loadings Stage 1",
             label = "tab:loadings",
             digits = 2),
      type = "latex",
      include.rownames = TRUE)



```

### AVE and composite reliability

```{r}
# Measurement-model reliability and convergent validity criteria:
# composite reliability and AVE follow Fornell & Larcker (1981),
# Bagozzi & Yi (1988), and Hair et al. (2017). Full references: REFERENCES.md.
# AVE
ave_values <- pls_model$outer_model %>%
  group_by(block) %>%
  summarize(AVE = sum((loading^2)) / (sum((loading^2)) + sum((1 - loading^2))))
# Composite Reliability
composite_reliability <- pls_model$outer_model %>%
  group_by(block) %>%
  summarize(CR = sum((loading)^2) / (sum((loading)^2) + sum((1 - loading)^2)))
print(ave_values)
print(composite_reliability)
```

### Fornell - Larcker criterion

```{r}
# Discriminant validity is assessed with the Fornell-Larcker criterion
# (Fornell & Larcker, 1981). Full reference: REFERENCES.md.
#AVE
ave_values <- pls_model$outer_model %>%
  group_by(block) %>%
  summarize(AVE = sum(loading^2) / length(loading))

#Fornell-Larcker Matrix
fornell_larcker <- matrix(NA, ncol=length(blocks), nrow=length(blocks))
rownames(fornell_larcker) <- colnames(fornell_larcker) <- c("Risk", "Attitudes", "PerceivedBehavC", "SubjectiveNorms", "Behaviour")

#Fill the diagonal with AVE values
diag(fornell_larcker) <- sqrt(ave_values$AVE)
latent_correlations <- cor(pls_model$scores)
# Latent correlations
for (i in 1:length(blocks)) {
  for (j in 1:length(blocks)) {
    if (i != j) {
      fornell_larcker[i, j] <- latent_correlations[i, j]
    }
  }
}
print(fornell_larcker)

fornell_larcker_df <- as.data.frame(fornell_larcker)
library(xtable)
# Creamos la tabla LaTeX
latex_table <- xtable(fornell_larcker,
                      caption = "Fornell-Larcker Matrix",
                      label = "tab:fornell_larcker",
                      digits = 2) 

xtable(fornell_larcker, caption = "Fornell-Larcker Criterion", label = "tab:FLMatrix")
```

### Heterotrait - monotrait Matrix

```{r}
# HTMT is used as an additional discriminant-validity criterion
# (Henseler, Ringle & Sarstedt, 2015). Full reference: REFERENCES.md.
library(semTools)
calculate_htmt <- function(pls_model, RHC, blocks) {
  # latent constructs
  scores <- pls_model$scores
  # HTMT matrix
  n_blocks <- length(blocks)
  htmt_matrix <- matrix(NA, nrow = n_blocks, ncol = n_blocks)
  
  
rownames(htmt_matrix) <- colnames(htmt_matrix) <- c("Risk", "Attitudes", "PerceivedBehavC", "SubjectiveNorms", "Behaviour")
  for (i in 1:(n_blocks - 1)) {
    for (j in (i + 1):n_blocks) {
      indicators_i <- blocks[[i]]
      indicators_j <- blocks[[j]]
      # HTMT correlations
      heterotrait_cor <- cor(RHC[, indicators_i], RHC[, indicators_j])
      monotrait_cor_i <- cor(RHC[, indicators_i])
      monotrait_cor_j <- cor(RHC[, indicators_j])
      # Calcular HTMT
      htmt_value <- mean(abs(heterotrait_cor)) / (sqrt(mean(abs(monotrait_cor_i[lower.tri(monotrait_cor_i)]))) * sqrt(mean(abs(monotrait_cor_j[lower.tri(monotrait_cor_j)]))))
            htmt_matrix[i, j] <- htmt_value
      htmt_matrix[j, i] <- htmt_value
    }
  }
  return(htmt_matrix)
}
htmt_matrix <- calculate_htmt(pls_model, RHC, blocks)
print(htmt_matrix)

xtable(htmt_matrix, caption = "Heterotrait-Monotrait Criterion", label = "tab:HTMTMatrix")
```

## Spanish firms sample


### Internal consistency
```{r}
psych::alpha(Spain[9:27])
```

### Confirmatory Factor Analysis

#### KMO and Fligner-Killeen test
```{r}
KMO((Spain[9:27]))
fligner.test((Spain[9:27]))
```

#### Factor loadings
```{r}
scree((Spain[9:27]), main="Sediment graph")
#Factor loadings
Behaviour <- psych::fa(Spain[9:27], nfactor = 3,rotate = "varimax", fm="pa") %>%
model_parameters(sort = TRUE, threshold = "max")

#Adter running the CFA, we select the cross loadings higher than .70
fac.SEM <- psych::fa(Spain[9:27], nfactor = 1,rotate = "none", fm="pa") %>% model_parameters(sort = TRUE, threshold = "max")
fac.SEM
```

### Measurement model
```{r}
path_matrix <- rbind(
  c(0, 0, 0, 0, 0), # Risk
  c(1, 0, 0, 0, 0), # Attitudes
  c(0, 1, 0, 0, 0), # PerceivedBehavC
  c(0, 0, 0, 0, 0), # SubjectiveNorms
  c(1, 1, 1, 1, 0)  # Behaviour
)
#first stage
colnames(path_matrix) <- rownames(path_matrix) <- c("Risk", "Attitudes", "PerceivedBehavC",   "SubjectiveNorms", "Behaviour")
# Measure model


#Stage1
blocks1 <- list(
  c("Rsk1","Rsk2", "Rsk3", "Rsk4"),       # Risk: Rsk1 presents a low load and a high cross loading with PBC, deletion.
  c("Att1", "Att2","Att3" , "Att4"),      # Attitudes:
  c("Pbc1", "Pbc2", "Pbc3", "Pbc4"),      # PerceivedBehavC: PBC2 has a higher load in Risk than in PBC, we delete it
  c("Sn1", "Sn2", "Sn3"),                 # SubjectiveNorms
  c("PastActionRed", "DecIntention", "PastOffset")          # Behaviour
)
#Stage2
blocks2 <- list(
  c("Rsk2","Rsk3","Rsk4"),          # Risk
  c("Att1", "Att2","Att3" ,"Att4"), # Attitudes: Att3 has a high cross loading in subjective norms 
  c( "Pbc2", "Pbc3", "Pbc4"),        # PerceivedBehavC: PBC1 presents low loading and  has a higher cross-load with subjective norms, we delete it
  c("Sn1", "Sn2", "Sn3"),           # SubjectiveNorms
  c("PastActionRed", "DecIntention")  # Behaviour: the construct presents an AVE under .50, so the lowest load is deleted to adjust composite reliability. 
)
#Stage3
blocks <- list(
  c("Rsk2", "Rsk3", "Rsk4"),
  c("Att1", "Att2", "Att4"),
  c( "Pbc3", "Pbc4"),
  c("Sn1", "Sn2", "Sn3"),
  c("PastActionRed", "DecIntention")
)

# Define all variables as reflexive
modes <- c("A", "A", "A", "A", "A")
#PLS-PM
pls_model <- plspm(Spain, path_matrix, blocks, modes)
summary(pls_model)



crossloadings <- pls_model$crossloadings  %>% dplyr::select( -block)

# Crear tabla LaTeX con xtable
xtable(crossloadings, caption = "Cross loadings PLS", label = "tab:loadings")




crossloadings_named <- crossloadings
rownames(crossloadings) <- crossloadings$name
crossloadings$name <- NULL
crossloadings$block <- NULL  # eliminar si no quieres mostrar 'block'

print(xtable(crossloadings,
             caption = "Cross loadings Stage 1",
             label = "tab:loadings",
             digits = 2),
      type = "latex",
      include.rownames = TRUE)

```



### AVE and composite reliability
```{r}
# Measurement-model reliability and convergent validity criteria:
# composite reliability and AVE follow Fornell & Larcker (1981),
# Bagozzi & Yi (1988), and Hair et al. (2017). Full references: REFERENCES.md.
# AVE
ave_values <- pls_model$outer_model %>%
  group_by(block) %>%
  summarize(AVE = sum((loading^2)) / (sum((loading^2)) + sum((1 - loading^2))))
# Composite Reliability
composite_reliability <- pls_model$outer_model %>%
  group_by(block) %>%
  summarize(CR = (sum(loading)^2) / ((sum(loading)^2) + sum(1 - loading^2)))
print(ave_values)
print(composite_reliability)
```

### Fornell - Larcker criterion
```{r}
# Discriminant validity is assessed with the Fornell-Larcker criterion
# (Fornell & Larcker, 1981). Full reference: REFERENCES.md.
library(dplyr)
#AVE
ave_values <- pls_model$outer_model %>%
  group_by(block) %>%
  summarize(AVE = sum(loading^2) / length(loading))

#Fornell-Larcker Matrix
fornell_larcker <- matrix(NA, ncol=length(blocks), nrow=length(blocks))
rownames(fornell_larcker) <- colnames(fornell_larcker) <- c("Risk", "Attitudes", "PerceivedBehavC",   "SubjectiveNorms", "Behaviour")

#Fill the diagonal with AVE values
diag(fornell_larcker) <- sqrt(ave_values$AVE)
latent_correlations <- cor(pls_model$scores)
# Latent correlations
for (i in 1:length(blocks)) {
  for (j in 1:length(blocks)) {
    if (i != j) {
      fornell_larcker[i, j] <- latent_correlations[i, j]
    }
  }
}
print(fornell_larcker)

fornell_larcker_df <- as.data.frame(fornell_larcker)
library(xtable)
# Creamos la tabla LaTeX
latex_table <- xtable(fornell_larcker,
                      caption = "Matriz de Fornell-Larcker",
                      label = "tab:fornell_larcker",
                      digits = 2)  # Ajusta la cantidad de decimales

xtable(fornell_larcker, caption = "Fornell-Larcker Criterion", label = "tab:FLMatrix")
```

### Heterotrait - monotrait Matrix
```{r}

# HTMT is used as an additional discriminant-validity criterion
# (Henseler, Ringle & Sarstedt, 2015). Full reference: REFERENCES.md.
library(semTools)
calculate_htmt <- function(pls_model, Spain, blocks) {
  # latent constructs
  scores <- pls_model$scores
  # HTMT matrix
  n_blocks <- length(blocks)
  htmt_matrix <- matrix(NA, nrow = n_blocks, ncol = n_blocks)
  rownames(htmt_matrix) <- colnames(htmt_matrix) <- c("Risk", "Attitudes", "PerceivedBehavC",   "SubjectiveNorms", "Behaviour")
  for (i in 1:(n_blocks - 1)) {
    for (j in (i + 1):n_blocks) {
      indicators_i <- blocks[[i]]
      indicators_j <- blocks[[j]]
      # HTMT correlations
      heterotrait_cor <- cor(Spain[, indicators_i], Spain[, indicators_j])
      monotrait_cor_i <- cor(Spain[, indicators_i])
      monotrait_cor_j <- cor(Spain[, indicators_j])
      # Calcular HTMT
      htmt_value <- mean(abs(heterotrait_cor)) / (sqrt(mean(abs(monotrait_cor_i[lower.tri(monotrait_cor_i)]))) * sqrt(mean(abs(monotrait_cor_j[lower.tri(monotrait_cor_j)]))))
            htmt_matrix[i, j] <- htmt_value
      htmt_matrix[j, i] <- htmt_value
    }
  }
  return(htmt_matrix)
}
htmt_matrix <- calculate_htmt(pls_model,Spain , blocks)
print(htmt_matrix)

xtable(htmt_matrix, caption = "Heterotrait-Monotrait Criterion", label = "tab:HTMTMatrix")
```






# Partial Least Squares - Structural Equation models
## Registry sample: Boot-strapping

```{r}
set.seed(123)
pls_model_boot <- plspm(RHC, path_matrix, blocks, modes, boot.val = TRUE, br = 10000
                        )
summary(pls_model_boot)
boot_results <- pls_model_boot$boot
print(boot_results)
```

### Latent scores' extraction for the logistic model

```{r}
library(writexl)
latent_scores <- pls_model_boot$scores
latent_scores_df <- as.data.frame(latent_scores)
SemConstructsE1 <- cbind(RHC, latent_scores_df)
file_path <- "SemConstructsE1.xlsx"
write_xlsx(SemConstructsE1, path = file_path)
```

## Spanish firms sample: Boot-strapping
```{r}
set.seed(0611)
library(plspm)
pls_model_boot_Spain <- plspm(Spain, path_matrix, blocks, modes, boot.val = TRUE, br = 10000)
summary(pls_model_boot_Spain)
```

### Latent scores' extraction for the logistic model

```{r}
library(writexl)
latent_scores <- pls_model_boot_Spain$scores
latent_scores_df <- as.data.frame(latent_scores)
SemConstructs <- cbind(Spain, latent_scores_df)
file_path <- "SemConstructsE2.xlsx"
write_xlsx(SemConstructs, path = file_path)
```


## Sensibility analysis for non-registered firms

```{r}
set.seed(0611)

# --- ORIGINAL MODEL (With Awareness in BEH) ---



path_matrix <- rbind(
  c(0, 0, 0, 0, 0), # Risk
  c(1, 0, 0, 0, 0), # Attitudes
  c(0, 1, 0, 0, 0), # PerceivedBehavC
  c(0, 0, 0, 0, 0), # SubjectiveNorms
  c(1, 1, 1, 1, 0)  # Behaviour
)
#first stage
colnames(path_matrix) <- rownames(path_matrix) <- c("Risk", "Attitudes", "PerceivedBehavC",   "SubjectiveNorms", "Behaviour")
# Measure model


#Stage3 straight
 blocks_original <- list(
  RSK = c("Rsk2", "Rsk3", "Rsk4"),     # Risk:
  ATT = c("Att1", "Att2", "Att4"),     # Attitudes:
  PBC = c("Pbc3", "Pbc4"),             # PerceivedBehavC:
  SN  = c("Sn1", "Sn2", "Sn3"),        # SubjectiveNorms
  BEH = c("PastActionRed", "DecIntention") # Behaviour
)

# Structural relation matrix (inner model)

modes <- rep("A", 5)  # reflective variables

model_original <- plspm(
  Spain,
  path_matrix = path_matrix,
  blocks = blocks_original,
  modes = modes,
  boot.val = TRUE,
  br = 10000
)

path_matrixsen <- rbind(
  c(0, 0, 0, 0, 0), # Risk
  c(1, 0, 0, 0, 0), # Attitudes
  c(0, 1, 0, 0, 0), # PerceivedBehavC
  c(0, 0, 0, 0, 0), # SubjectiveNorms
  c(1, 1, 1, 1, 0)  # Behaviour
)


colnames(path_matrixsen) <- rownames(path_matrixsen) <- c("Risk", "Attitudes", "PerceivedBehavC",   "SubjectiveNorms", "Behaviour")
# Measure model
# --- Sensibility model (without Awareness) ---
blocks_sensitivity <- list(
  RSK = c("Rsk2", "Rsk3", "Rsk4"),     # Risk:
  ATT = c("Att1", "Att2", "Att4"),     # Attitudes:
  PBC = c("Pbc3", "Pbc4"),             # PerceivedBehavC:
  SN  = c("Sn1", "Sn2", "Sn3"),        # SubjectiveNorms
  BEH = c("DecIntention") # Behaviour
)
modessen <- rep("A", 5)

model_sensitivity <- plspm(
  Spain,
  path_matrix = path_matrixsen,
  blocks = blocks_sensitivity,
  modes = modessen,
  boot.val = TRUE,
  br = 10000
)

summary(pls_model_boot_Spain)
summary(model_sensitivity)

```


# Random utility Framework

## Data
```{r}
RHC <- read_excel(data_path, sheet = "DCE_RHC_II")
ESP <- read_excel(data_path, sheet = "DCE_Spain_II")

# --- 1)Atributes
attr_Project  <- intersect(c("Project", "project"), names(RHC))[1]
attr_Location <- intersect(c("Location", "location"), names(RHC))[1]
attr_Species  <- intersect(c("Species", "Specie", "species", "specie"), names(RHC))[1]
attr_Price    <- intersect(c("Price", "price"), names(RHC))[1]

stopifnot(!is.na(attr_Project), !is.na(attr_Location),
          !is.na(attr_Species), !is.na(attr_Price))

attrs <- c(attr_Project, attr_Location, attr_Species, attr_Price)

# --- 2) Create Alt (optout/project) and ASC_accept from ASC_optout
RHC <- RHC %>%
  mutate(
    Alt = ifelse(Optout == 1, "optout", "project"),
    ASC_accept = ifelse(Optout == 1, 0, 1)
  )

# --- 3) Replace -4 by 0 only in opt-out rows
RHC <- RHC %>%
  mutate(across(all_of(attrs), ~ ifelse(Alt == "optout" & .x == -4, 0, .x)))

# --- 4) Create chid (ID by choice card)
has_qid  <- "QuestionId" %in% names(RHC) || "Question_Id" %in% names(RHC)
qid_name <- if ("QuestionId" %in% names(RHC)) "QuestionId" else "Question_Id"
id_name  <- if ("Id" %in% names(RHC)) "Id" else "id"
pqi_name <- if ("ParentQuestionId" %in% names(RHC)) "ParentQuestionId" else "Parent_Question_Id"

if (has_qid) {
  RHC <- RHC %>%
    mutate(chid = paste(.data[[id_name]], .data[[pqi_name]], .data[[qid_name]], sep = "_"))
} else {
  ord_name <- if ("Ordering" %in% names(RHC)) "Ordering" else stop("Falta QuestionId y Ordering")
  RHC <- RHC %>%
    mutate(chid = paste(.data[[id_name]], .data[[pqi_name]], .data[[ord_name]], sep = "_"))
}

# --- 5) Verification (Each card must have only one choice)
RHC %>%
  group_by(chid) %>%
  summarise(chosen_sum = sum(Choice), .groups = "drop") %>%
  filter(chosen_sum != 1)

# --- 6) mlogit definition
library(mlogit)
df_mld_RHC <- mlogit.data(RHC,
                           choice   = "Choice",
                           shape    = "long",
                           alt.var  = "Alt",
                           chid.var = "chid",
                           id.var   = "Id")


# --- 1)Atributes
attr_Location <- intersect(c("Location", "location"), names(ESP))[1]
attr_Species  <- intersect(c("Species", "Specie", "species", "specie"), names(ESP))[1]
attr_Price    <- intersect(c("Price", "price"), names(ESP))[1]

stopifnot(!is.na(attr_Project), !is.na(attr_Location),
          !is.na(attr_Species), !is.na(attr_Price))

attrs <- c(attr_Project, attr_Location, attr_Species, attr_Price)

# --- 2) Create Alt (optout/project) and ASC_accept from ASC_optout
ESP <- ESP %>%
  mutate(
    Alt = ifelse(Optout == 1, "optout", "project"),
    ASC_accept = ifelse(Optout == 1, 0, 1)
  )

# --- 3) Replace -4 by 0 only in opt-out rows
ESP <- ESP %>%
  mutate(across(all_of(attrs), ~ ifelse(Alt == "optout" & .x == -4, 0, .x)))

# --- 4) Create chid (ID by choice card)
has_qid  <- "QuestionId" %in% names(ESP) || "Question_Id" %in% names(ESP)
qid_name <- if ("QuestionId" %in% names(ESP)) "QuestionId" else "Question_Id"
id_name  <- if ("Id" %in% names(ESP)) "Id" else "id"
pqi_name <- if ("ParentQuestionId" %in% names(ESP)) "ParentQuestionId" else "Parent_Question_Id"

if (has_qid) {
  ESP <- ESP %>%
    mutate(chid = paste(.data[[id_name]], .data[[pqi_name]], .data[[qid_name]], sep = "_"))
} else {
  ord_name <- if ("Ordering" %in% names(ESP)) "Ordering" else stop("There are no variables named QuestionId and/or Ordering")
  ESP <- ESP %>%
    mutate(chid = paste(.data[[id_name]], .data[[pqi_name]], .data[[ord_name]], sep = "_"))
}

# --- 5) Verification (Each card must have only one choice)
ESP %>%
  group_by(chid) %>%
  summarise(chosen_sum = sum(Choice), .groups = "drop") %>%
  filter(chosen_sum != 1)

# --- 6) mlogit definition
library(mlogit)
df_mld_ESP <- mlogit.data(ESP,
                           choice   = "Choice",
                           shape    = "long",
                           alt.var  = "Alt",
                           chid.var = "chid",
                           id.var   = "Id")
```


## Conditional variables interaction with ASC
```{r}
df_mld_RHC$cRisk  <- scale(df_mld_RHC$Risk, center=TRUE, scale=FALSE)
df_mld_RHC$cAtt <- scale(df_mld_RHC$Attitudes, center=TRUE, scale=FALSE)
df_mld_RHC$cSN  <- scale(df_mld_RHC$SubjectiveNorms, center=TRUE, scale=FALSE)
df_mld_RHC$cPBC  <- scale(df_mld_RHC$PerceivedBehavC, center=TRUE, scale=FALSE)

df_mld_RHC$ASC_Att <- df_mld_RHC$ASC_accept * df_mld_RHC$cAtt
df_mld_RHC$ASC_SN  <- df_mld_RHC$ASC_accept * df_mld_RHC$cSN
df_mld_RHC$ASC_PBC <- df_mld_RHC$ASC_accept * df_mld_RHC$cPBC 
df_mld_RHC$ASC_Risk <- df_mld_RHC$ASC_accept * df_mld_RHC$cRisk
df_mld_RHC$ASC_Service <- df_mld_RHC$ASC_accept * df_mld_RHC$Service
df_mld_RHC$ASC_PYME <- df_mld_RHC$ASC_accept * df_mld_RHC$PYME
df_mld_RHC$ASC_SustDepart <- df_mld_RHC$ASC_accept * df_mld_RHC$SustDepart

df_mld_ESP$cRisk  <- scale(df_mld_ESP$Risk, center=TRUE, scale=FALSE)
df_mld_ESP$cAtt <- scale(df_mld_ESP$Attitudes, center=TRUE, scale=FALSE)
df_mld_ESP$cSN  <- scale(df_mld_ESP$SubjectiveNorms, center=TRUE, scale=FALSE)
df_mld_ESP$cPBC  <- scale(df_mld_ESP$PerceivedBehavC, center=TRUE, scale=FALSE)

df_mld_ESP$ASC_Att <- df_mld_ESP$ASC_accept * df_mld_ESP$cAtt
df_mld_ESP$ASC_SN  <- df_mld_ESP$ASC_accept * df_mld_ESP$cSN
df_mld_ESP$ASC_PBC <- df_mld_ESP$ASC_accept * df_mld_ESP$cPBC
df_mld_ESP$ASC_Risk <- df_mld_ESP$ASC_accept * df_mld_ESP$cRisk
df_mld_ESP$ASC_Service <- df_mld_ESP$ASC_accept * df_mld_ESP$Service
df_mld_ESP$ASC_PYME <- df_mld_ESP$ASC_accept * df_mld_ESP$PYME
df_mld_ESP$ASC_SustDepart <- df_mld_ESP$ASC_accept * df_mld_ESP$SustDepart
```
## Conditional models

### Base multinomial logit
```{r}


# MNL baseline (atributes + ASC) ---
mnlR0 <- mlogit(Choice ~ 0 + Project + Specie + Location + Price  + ASC_accept,
               data = df_mld_RHC)
summary(mnlR0)

# MNL baseline (atributes + ASC) ---
mnlE0 <- mlogit(Choice ~ 0 + Project + Specie + Location + Price + ASC_accept,
               data = df_mld_ESP)
summary(mnlE0)
```

### Conditional logit: Registry
```{r}
clRec <- gmnl(
  Choice ~ 0 + Project + Specie + Location + Price + ASC_accept + ASC_Att + ASC_SN + ASC_PBC+ ASC_Risk + ASC_Service + ASC_PYME,
  data = df_mld_RHC,
  model = "mnl"
)
summary(clRec)

```
### Conditional logit: Spain
```{r}
clEec <- gmnl(
  Choice ~ 0 + Project + Specie + Location + Price + ASC_accept + ASC_Att + ASC_SN + ASC_PBC+ ASC_Risk + ASC_Service + ASC_PYME,
  data = df_mld_ESP,
  model = "mnl"
)
summary(clEec)

```

### Mixed logit: Halton draws

#### Functions' definition 
```{r}
set.seed(45855)

# ===============================
# ROBUST AUXILIAR FUNCTIONS
# ===============================
len1 <- function(x){
  if (is.null(x) || length(x) == 0) return(NA_real_)
  y <- suppressWarnings(as.numeric(x))
  if (length(y) == 0) return(NA_real_)
  if (!is.finite(y[1])) return(NA_real_) else return(y[1])
}

extract_se <- function(m){
  cf <- coef(m)
  out <- tryCatch({
    V  <- vcov(m)
    se <- sqrt(diag(V))
    if(is.null(names(se)) && !is.null(rownames(V))) names(se) <- rownames(V)
    if(is.null(names(se))) names(se) <- names(cf)
    se
  }, error = function(e) {
    se <- rep(NA_real_, length(cf))
    names(se) <- names(cf)
    se
  })
  out
}

compare_models_hybrid <- function(m_new, m_old, beta_switch, se_switch){
  b_new <- coef(m_new); b_old <- coef(m_old)
  alln  <- union(names(b_new), names(b_old))
  b_new <- b_new[alln]; b_old <- b_old[alln]

  se_new <- extract_se(m_new)[alln]
  se_old <- extract_se(m_old)[alln]

  db_abs  <- abs(b_new - b_old)
  db_rel  <- abs(b_new - b_old) / (abs(b_old) + 1e-12)
  dse_abs <- abs(se_new - se_old)
  dse_rel <- abs(se_new - se_old) / (abs(se_old) + 1e-12)

  beta_metric <- ifelse(abs(b_old) < beta_switch, db_abs, db_rel)
  se_metric   <- ifelse(abs(se_old) < se_switch, dse_abs, dse_rel)

  i_beta <- which.max(ifelse(is.finite(beta_metric), beta_metric, -Inf))
  i_se   <- which.max(ifelse(is.finite(se_metric),   se_metric,   -Inf))

  top_beta_name <- if(length(i_beta)) names(beta_metric)[i_beta] else NA_character_
  top_se_name   <- if(length(i_se))   names(se_metric)[i_se]      else NA_character_

  list(
    beta_metric = beta_metric,
    se_metric   = se_metric,
    top_beta    = list(name = top_beta_name,
                       old  = len1(b_old[top_beta_name]),
                       new  = len1(b_new[top_beta_name]),
                       met  = len1(beta_metric[top_beta_name]),
                       used = ifelse(abs(len1(b_old[top_beta_name])) < beta_switch, "abs", "rel")),
    top_se      = list(name = top_se_name,
                       old  = len1(se_old[top_se_name]),
                       new  = len1(se_new[top_se_name]),
                       met  = len1(se_metric[top_se_name]),
                       used = ifelse(abs(len1(se_old[top_se_name])) < se_switch, "abs", "rel")),
    max_beta = suppressWarnings(max(beta_metric[is.finite(beta_metric)], na.rm=TRUE)),
    max_se   = suppressWarnings(max(se_metric[is.finite(se_metric)],     na.rm=TRUE)),
    dll      = abs(len1(logLik(m_new)) - len1(logLik(m_old)))
  )
}
```
#### Registry Halton draws loop halton = list(c(2, 3, 5, 7)) BGFS method

```{r}
set.seed(1685)
# ===============================
# 1) SPECIFICATION
# ===============================
form_mixlR2 <- Choice ~ 0 + Project + Specie + Location + Price +
  ASC_accept + ASC_Att + ASC_SN + ASC_PBC + ASC_PYME + ASC_Service

ranp <- c(
  Project    = "n",
  Specie     = "n",
  Location   = "n",
  ASC_accept = "n"
)

data <- df_mld_RHC

# ===============================
# 2) LOOP PARAMETERS
# ===============================
R_start <- 2000
step_R  <- 200
max_R   <- 3000

# ---------------------------------------------------------------------------
# Tolerance thresholds — grounded in discrete choice literature.
# Full APA 6th references are listed separately in REFERENCES.md.
#
# beta_switch / se_switch:
#   Below this |value|, use ABSOLUTE metric; above it, use RELATIVE metric.
#   Set to 0.05: parameters near zero are essentially null and a relative
#   criterion would be artificially loose (Train, 2009).
#
# Mean parameters (beta_abs/rel_tol_mean):
#   0.001 (0.1% relative, or 0.001 absolute).
#   Hensher & Greene (2003) "State of Practice": stability declared when
#   parameters do not change in the third significant digit across draw counts.
#   Train (2009): simulation error should be negligible relative to the
#   sampling variance of the estimator.
#
# SD parameters (beta_abs/rel_tol_sd):
#   Loosened to 0.005 (0.5%). Standard deviations of mixing distributions
#   carry inherently higher simulation noise than means; Czajkowski &
#   Budzinski (2019) set their Minimum Tolerance
#   criterion at this order of magnitude for SDs estimated with Halton draws.
#
# SE tolerance (se_abs/rel_tol):
#   Hess, Train & Polak (2006): simulation error in SEs
#   should be evaluated relative to the SE magnitude itself.
#   2% relative (0.02) is consistent with their Monte Carlo comparisons.
#   Absolute fallback of 0.01 for near-zero SEs.
#
# Log-likelihood relative change (ll_rel_tol):
#   1e-5. Train (2009): MSL is asymptotically equivalent to ML when R grows
#   faster than sqrt(N); at that point ΔLL between consecutive R values
#   should be negligible. 1e-5 is stricter than the typical optimizer
#   convergence tolerance (1e-4) and consistent with Czajkowski & Budzinski
#   (2019) simulation error benchmarks.
# ---------------------------------------------------------------------------
beta_switch        <- 0.05   # |b_old| threshold: abs vs. rel metric

beta_abs_tol_mean  <- 0.001  # abs Δ tolerance for mean parameters
beta_rel_tol_mean  <- 0.001  # rel Δ tolerance for mean parameters

beta_abs_tol_sd    <- 0.005  # abs Δ tolerance for sd. parameters
beta_rel_tol_sd    <- 0.005  # rel Δ tolerance for sd. parameters

se_switch          <- 0.05   # SE threshold: abs vs. rel metric
se_abs_tol         <- 0.010  # abs Δ tolerance for standard errors
se_rel_tol         <- 0.020  # rel Δ tolerance for standard errors (2%)

ll_rel_tol         <- 1e-5   # rel ΔLL tolerance across consecutive R values

# ===============================
# 3) SAFE WRAPPER CON BFGS
# ===============================
fit_gmnl_safe <- function(formula, data, ranp, R, corr = FALSE, start = NULL){
  gmnl(formula, data=data, model="mixl", ranp=ranp, panel=TRUE,
       R=R, halton=list(prime = c(11, 3, 5, 7)), correlation=corr,
       method="bfgs", start=start)
}

# ===============================
# 4) Stabilisation loop
# ===============================
fits_A <- list()
diag_tbl_R <- data.frame(
  R = integer(), logLik = numeric(),
  max_beta_metric = numeric(), mover_beta = character(), metric_used_beta = character(),
  max_se_metric   = numeric(), mover_se   = character(), metric_used_se   = character(),
  dLL = numeric(), dLL_rel = numeric(),
  stringsAsFactors = FALSE
)

prev_fit   <- NULL
prev_coefs <- NULL
Rcurr <- as.integer(R_start)

repeat{
  cat(sprintf("\n[gmnl] MIXL-R (corr=FALSE) con R = %d (Halton, BFGS) ...\n", Rcurr))
  t0 <- Sys.time()
  fitA <- fit_gmnl_safe(form_mixlR2, data = data, ranp = ranp,
                        R = Rcurr, corr = FALSE, start = prev_coefs)
  t1 <- Sys.time()
  llA <- len1(logLik(fitA))
  cat(sprintf("  Hecho en %.1f s | logLik = %.3f\n",
              as.numeric(difftime(t1, t0, units="secs")), llA))

  fits_A[[as.character(Rcurr)]] <- fitA
  coefs_now <- tryCatch(coef(fitA), error=function(e) NULL)

  if(is.null(prev_fit)){
    diag_tbl_R <- rbind(diag_tbl_R, data.frame(
      R = Rcurr, logLik = llA,
      max_beta_metric = NA_real_, mover_beta = NA_character_, metric_used_beta = NA_character_,
      max_se_metric   = NA_real_, mover_se   = NA_character_, metric_used_se   = NA_character_,
      dLL = NA_real_, dLL_rel = NA_real_
    ))
  } else {
    cmp <- compare_models_hybrid(fitA, prev_fit, beta_switch, se_switch)
    dLL_abs <- len1(cmp$dll)
    dLL_rel <- abs(dLL_abs) / (abs(len1(logLik(prev_fit))) + 1e-12)

    cat(sprintf("  Summing up: max β-metric=%.6f | max SE-metric=%.6f | ΔLL_rel=%.6e\n",
                len1(cmp$max_beta), len1(cmp$max_se), dLL_rel))

    row <- data.frame(
      R = Rcurr, logLik = llA,
      max_beta_metric = len1(cmp$max_beta),
      mover_beta      = cmp$top_beta$name,
      metric_used_beta= cmp$top_beta$used,
      max_se_metric   = len1(cmp$max_se),
      mover_se        = cmp$top_se$name,
      metric_used_se  = cmp$top_se$used,
      dLL             = dLL_abs,
      dLL_rel         = dLL_rel,
      stringsAsFactors = FALSE
    )

    # ---- MEAN VS. SD separate criteria ----
    b_new <- coef(fitA); b_old <- coef(prev_fit)
    alln  <- union(names(b_new), names(b_old))
    b_new <- b_new[alln]; b_old <- b_old[alln]
    is_sd <- grepl("^sd\\.", alln)

    db_abs <- abs(b_new - b_old)
    db_rel <- abs(b_new - b_old) / (abs(b_old) + 1e-12)

    # Mean
    db_abs_mean <- db_abs[!is_sd]; db_rel_mean <- db_rel[!is_sd]; b_old_mean <- b_old[!is_sd]
    beta_ok_mean <- all(ifelse(abs(b_old_mean) < beta_switch,
                               db_abs_mean < beta_abs_tol_mean,
                               db_rel_mean < beta_rel_tol_mean), na.rm = TRUE)

    # Deviations
    db_abs_sd <- db_abs[is_sd]; db_rel_sd <- db_rel[is_sd]; b_old_sd <- b_old[is_sd]
    beta_ok_sd <- all(ifelse(abs(b_old_sd) < beta_switch,
                             db_abs_sd < beta_abs_tol_sd,
                             db_rel_sd < beta_rel_tol_sd), na.rm = TRUE)

    beta_ok <- beta_ok_mean && beta_ok_sd
    se_ok   <- ifelse(row$metric_used_se == "rel",
                      row$max_se_metric < se_rel_tol,
                      row$max_se_metric < se_abs_tol)
    ll_ok   <- row$dLL_rel < ll_rel_tol

    diag_tbl_R <- rbind(diag_tbl_R, row)

    # ---- Stopping conditions ----
    if (!is.na(beta_ok) && !is.na(se_ok) && !is.na(ll_ok) && beta_ok && se_ok && ll_ok) {
      cat(">> The model reach stability (mean, sd, SE y ΔLL_rel). Stop.\n")
      prev_fit   <- fitA
      prev_coefs <- coefs_now
      break
    }
  }

  # ---- UPPER LIMIT CONTROL ----
  if (Rcurr + step_R > max_R){
    cat(sprintf(">> The model reach max_R=%d without reaching tolerance criteria. Stop.\n", max_R))
    prev_fit   <- fitA
    prev_coefs <- coefs_now
    break
  }

  prev_fit   <- fitA
  prev_coefs <- coefs_now
  Rcurr      <- Rcurr + step_R
}

# ===============================
# 5) Final model
# ===============================
R_final_A   <- diag_tbl_R$R[nrow(diag_tbl_R)]
mixlRHC <- fits_A[[as.character(R_final_A)]]

cat(sprintf("\n[MIXL-A final)] R = %s | logLik = %.3f\n",
            R_final_A, len1(logLik(mixlRHC))))
summary(mixlRHC)

# Iteration's diagnose
diag_tbl_R

```


#### Reduced Mixed logit 

```{r}

#set.seed(1685) semilla del modelo reportado en el manuscrito, no borrar
set.seed(0611) #semilla para comprobar la estabilidad del modelo, misma significación, modelo y especificaicones correctas

# ===============================
# 1) SPECIFICATION: MIXL REDUCED RHC
# Project and Specie remain in utility, but are treated as fixed parameters
# Random parameters: Location and ASC_accept only
# ===============================

form_mixl_redRHC <- Choice ~ 0 + Project + Specie + Location + Price +
  ASC_accept + ASC_Att + ASC_SN + ASC_PBC + ASC_PYME + ASC_Service

ranp_redRHC <- c(
  Location   = "n",
  ASC_accept = "n"
)

data_redRHC <- df_mld_RHC

# ===============================
# 2) LOOP PARAMETERS
# ===============================

R_start_redRHC <- 3400
step_R_redRHC  <- 200
max_R_redRHC   <- 4000

# ---------------------------------------------------------------------------
# Tolerance thresholds — same criteria as the full RHC mixed logit loop.
# Full APA 6th references are listed separately in REFERENCES.md.
#
# beta_switch / se_switch = 0.05: use absolute differences near zero and
# relative differences otherwise (Train, 2009).
# Mean parameters = 0.001: stability in approximately the third significant
# digit across draw counts (Hensher & Greene, 2003; Train, 2009).
# SD parameters = 0.005: looser criterion for mixing-distribution standard
# deviations, which are more simulation-noisy (Czajkowski & Budzinski, 2019).
# SE tolerance = 0.010 absolute / 0.020 relative: evaluate simulation error
# relative to the SE magnitude itself (Hess, Train & Polak, 2006).
# Log-likelihood tolerance = 1e-5: negligible relative change across
# consecutive draw counts (Train, 2009; Czajkowski & Budzinski, 2019).
# ---------------------------------------------------------------------------

beta_switch_redRHC        <- 0.05

beta_abs_tol_mean_redRHC  <- 0.001
beta_rel_tol_mean_redRHC  <- 0.001

beta_abs_tol_sd_redRHC    <- 0.005
beta_rel_tol_sd_redRHC    <- 0.005

se_switch_redRHC          <- 0.05
se_abs_tol_redRHC         <- 0.010
se_rel_tol_redRHC         <- 0.020

ll_rel_tol_redRHC         <- 1e-5

# ===============================
# 3) SAFE WRAPPER WITH BFGS
# ===============================

fit_gmnl_safe_redRHC <- function(formula, data, ranp, R, corr = FALSE, start = NULL){
  gmnl(
    formula,
    data = data,
    model = "mixl",
    ranp = ranp,
    panel = TRUE,
    R = R,
    halton = list(prime = c(2, 3)),
    correlation = corr,
    method = "bfgs",
    start = start
  )
}

# ===============================
# 4) STABILISATION LOOP
# ===============================

fits_redRHC <- list()

diag_tbl_R_redRHC <- data.frame(
  R = integer(),
  logLik = numeric(),
  max_beta_metric = numeric(),
  mover_beta = character(),
  metric_used_beta = character(),
  max_se_metric = numeric(),
  mover_se = character(),
  metric_used_se = character(),
  dLL = numeric(),
  dLL_rel = numeric(),
  stringsAsFactors = FALSE
)

prev_fit_redRHC   <- NULL
prev_coefs_redRHC <- NULL
Rcurr_redRHC      <- as.integer(R_start_redRHC)

repeat{

  cat(sprintf(
    "\n[gmnl] MIXL-RHC reduced (corr=FALSE) con R = %d (Halton, BFGS) ...\n",
    Rcurr_redRHC
  ))

  t0_redRHC <- Sys.time()

  fit_redRHC_iter <- fit_gmnl_safe_redRHC(
    formula = form_mixl_redRHC,
    data    = data_redRHC,
    ranp    = ranp_redRHC,
    R       = Rcurr_redRHC,
    corr    = FALSE,
    start   = prev_coefs_redRHC
  )

  t1_redRHC <- Sys.time()

  ll_redRHC_iter <- len1(logLik(fit_redRHC_iter))

  cat(sprintf(
    "  Hecho en %.1f s | logLik = %.3f\n",
    as.numeric(difftime(t1_redRHC, t0_redRHC, units = "secs")),
    ll_redRHC_iter
  ))

  fits_redRHC[[as.character(Rcurr_redRHC)]] <- fit_redRHC_iter

  coefs_now_redRHC <- tryCatch(
    coef(fit_redRHC_iter),
    error = function(e) NULL
  )

  if(is.null(prev_fit_redRHC)){

    diag_tbl_R_redRHC <- rbind(
      diag_tbl_R_redRHC,
      data.frame(
        R = Rcurr_redRHC,
        logLik = ll_redRHC_iter,
        max_beta_metric = NA_real_,
        mover_beta = NA_character_,
        metric_used_beta = NA_character_,
        max_se_metric = NA_real_,
        mover_se = NA_character_,
        metric_used_se = NA_character_,
        dLL = NA_real_,
        dLL_rel = NA_real_,
        stringsAsFactors = FALSE
      )
    )

  } else {

    cmp_redRHC <- compare_models_hybrid(
      fit_redRHC_iter,
      prev_fit_redRHC,
      beta_switch_redRHC,
      se_switch_redRHC
    )

    dLL_abs_redRHC <- len1(cmp_redRHC$dll)

    dLL_rel_redRHC <- abs(dLL_abs_redRHC) /
      (abs(len1(logLik(prev_fit_redRHC))) + 1e-12)

    cat(sprintf(
      "  Summing up: max β-metric=%.6f | max SE-metric=%.6f | ΔLL_rel=%.6e\n",
      len1(cmp_redRHC$max_beta),
      len1(cmp_redRHC$max_se),
      dLL_rel_redRHC
    ))

    row_redRHC <- data.frame(
      R = Rcurr_redRHC,
      logLik = ll_redRHC_iter,
      max_beta_metric = len1(cmp_redRHC$max_beta),
      mover_beta = cmp_redRHC$top_beta$name,
      metric_used_beta = cmp_redRHC$top_beta$used,
      max_se_metric = len1(cmp_redRHC$max_se),
      mover_se = cmp_redRHC$top_se$name,
      metric_used_se = cmp_redRHC$top_se$used,
      dLL = dLL_abs_redRHC,
      dLL_rel = dLL_rel_redRHC,
      stringsAsFactors = FALSE
    )

    # ---- MEAN VS. SD separate criteria ----

    b_new_redRHC <- coef(fit_redRHC_iter)
    b_old_redRHC <- coef(prev_fit_redRHC)

    alln_redRHC <- union(names(b_new_redRHC), names(b_old_redRHC))

    b_new_redRHC <- b_new_redRHC[alln_redRHC]
    b_old_redRHC <- b_old_redRHC[alln_redRHC]

    is_sd_redRHC <- grepl("^sd\\.", alln_redRHC)

    db_abs_redRHC <- abs(b_new_redRHC - b_old_redRHC)

    db_rel_redRHC <- abs(b_new_redRHC - b_old_redRHC) /
      (abs(b_old_redRHC) + 1e-12)

    # Mean parameters

    db_abs_mean_redRHC <- db_abs_redRHC[!is_sd_redRHC]
    db_rel_mean_redRHC <- db_rel_redRHC[!is_sd_redRHC]
    b_old_mean_redRHC  <- b_old_redRHC[!is_sd_redRHC]

    beta_ok_mean_redRHC <- all(
      ifelse(
        abs(b_old_mean_redRHC) < beta_switch_redRHC,
        db_abs_mean_redRHC < beta_abs_tol_mean_redRHC,
        db_rel_mean_redRHC < beta_rel_tol_mean_redRHC
      ),
      na.rm = TRUE
    )

    # Standard deviation parameters

    db_abs_sd_redRHC <- db_abs_redRHC[is_sd_redRHC]
    db_rel_sd_redRHC <- db_rel_redRHC[is_sd_redRHC]
    b_old_sd_redRHC  <- b_old_redRHC[is_sd_redRHC]

    beta_ok_sd_redRHC <- all(
      ifelse(
        abs(b_old_sd_redRHC) < beta_switch_redRHC,
        db_abs_sd_redRHC < beta_abs_tol_sd_redRHC,
        db_rel_sd_redRHC < beta_rel_tol_sd_redRHC
      ),
      na.rm = TRUE
    )

    beta_ok_redRHC <- beta_ok_mean_redRHC && beta_ok_sd_redRHC

    se_ok_redRHC <- ifelse(
      row_redRHC$metric_used_se == "rel",
      row_redRHC$max_se_metric < se_rel_tol_redRHC,
      row_redRHC$max_se_metric < se_abs_tol_redRHC
    )

    ll_ok_redRHC <- row_redRHC$dLL_rel < ll_rel_tol_redRHC

    diag_tbl_R_redRHC <- rbind(diag_tbl_R_redRHC, row_redRHC)

    # ---- Stopping conditions ----

    if (
      !is.na(beta_ok_redRHC) &&
      !is.na(se_ok_redRHC) &&
      !is.na(ll_ok_redRHC) &&
      beta_ok_redRHC &&
      se_ok_redRHC &&
      ll_ok_redRHC
    ) {
      cat(">> The reduced MIXL-RHC model reached stability (mean, sd, SE and ΔLL_rel). Stop.\n")

      prev_fit_redRHC   <- fit_redRHC_iter
      prev_coefs_redRHC <- coefs_now_redRHC

      break
    }
  }

  # ---- UPPER LIMIT CONTROL ----

  if (Rcurr_redRHC + step_R_redRHC > max_R_redRHC){

    cat(sprintf(
      ">> The reduced MIXL-RHC model reached max_R=%d without reaching all tolerance criteria. Stop.\n",
      max_R_redRHC
    ))

    prev_fit_redRHC   <- fit_redRHC_iter
    prev_coefs_redRHC <- coefs_now_redRHC

    break
  }

  prev_fit_redRHC   <- fit_redRHC_iter
  prev_coefs_redRHC <- coefs_now_redRHC
  Rcurr_redRHC      <- Rcurr_redRHC + step_R_redRHC
}

# ===============================
# 5) FINAL MODEL: REDUCED MIXL RHC
# ===============================

R_final_redRHC <- diag_tbl_R_redRHC$R[nrow(diag_tbl_R_redRHC)]

mixlRHC_red <- fits_redRHC[[as.character(R_final_redRHC)]]

cat(sprintf(
  "\n[MIXL-RHC reduced final] R = %s | logLik = %.3f\n",
  R_final_redRHC,
  len1(logLik(mixlRHC_red))
))

summary(mixlRHC_red)

# Iteration diagnosis
diag_tbl_R_redRHC

```




### Spain Halton draws loop halton = list(c(2, 11, 7, 13)) BGFS method
```{r}

 set.seed(1685)
# ===============================
# 1) SPECIFICATION
# ===============================
form_mixlE2 <- Choice ~ 0 + Project + Specie + Location + Price +
  ASC_accept + ASC_Att + ASC_SN + ASC_PBC + ASC_PYME + ASC_Service

ranp <- c(
  Project    = "n",
  Specie     = "n",
  Location   = "n",
  ASC_accept = "n"
)

data <- df_mld_ESP

# ===============================
# 2) LOOP PARAMETERS
# ===============================
R_start <- 2000
step_R  <- 200
max_R   <- 8000

# ---------------------------------------------------------------------------
# Tolerance thresholds — grounded in discrete choice literature.
# Full APA 6th references are listed separately in REFERENCES.md.
#
# beta_switch / se_switch:
#   Below this |value|, use ABSOLUTE metric; above it, use RELATIVE metric.
#   Set to 0.05: parameters near zero are essentially null and a relative
#   criterion would be artificially loose (Train, 2009).
#
# Mean parameters (beta_abs/rel_tol_mean):
#   0.001 (0.1% relative, or 0.001 absolute).
#   Hensher & Greene (2003) "State of Practice": stability declared when
#   parameters do not change in the third significant digit across draw counts.
#   Train (2009): simulation error should be negligible relative to the
#   sampling variance of the estimator.
#
# SD parameters (beta_abs/rel_tol_sd):
#   Loosened to 0.005 (0.5%). Standard deviations of mixing distributions
#   carry inherently higher simulation noise than means; Czajkowski &
#   Budzinski (2019) set their Minimum Tolerance
#   criterion at this order of magnitude for SDs estimated with Halton draws.
#
# SE tolerance (se_abs/rel_tol):
#   Hess, Train & Polak (2006): simulation error in SEs
#   should be evaluated relative to the SE magnitude itself.
#   2% relative (0.02) is consistent with their Monte Carlo comparisons.
#   Absolute fallback of 0.01 for near-zero SEs.
#
# Log-likelihood relative change (ll_rel_tol):
#   1e-5. Train (2009): MSL is asymptotically equivalent to ML when R grows
#   faster than sqrt(N); at that point ΔLL between consecutive R values
#   should be negligible. 1e-5 is stricter than the typical optimizer
#   convergence tolerance (1e-4) and consistent with Czajkowski & Budzinski
#   (2019) simulation error benchmarks.
# ---------------------------------------------------------------------------
beta_switch        <- 0.05   # |b_old| threshold: abs vs. rel metric

beta_abs_tol_mean  <- 0.001  # abs Δ tolerance for mean parameters
beta_rel_tol_mean  <- 0.001  # rel Δ tolerance for mean parameters

beta_abs_tol_sd    <- 0.005  # abs Δ tolerance for sd. parameters
beta_rel_tol_sd    <- 0.005  # rel Δ tolerance for sd. parameters

se_switch          <- 0.05   # SE threshold: abs vs. rel metric
se_abs_tol         <- 0.010  # abs Δ tolerance for standard errors
se_rel_tol         <- 0.020  # rel Δ tolerance for standard errors (2%)

ll_rel_tol         <- 1e-5   # rel ΔLL tolerance across consecutive R values

# ===============================
# 3) SAFE WRAPPER CON BFGS
# ===============================
fit_gmnl_safe <- function(formula, data, ranp, R, corr = FALSE, start = NULL){
  gmnl(formula, data=data, model="mixl", ranp=ranp, panel=TRUE,
       R=R, halton=list(prime = c(2, 3, 5, 7)), correlation=corr,
       method="bfgs", start=start)
}

# ===============================
# 4) Stabilisation loop
# ===============================
fits_E <- list()
diag_tbl_E <- data.frame(
  R = integer(), logLik = numeric(),
  max_beta_metric = numeric(), mover_beta = character(), metric_used_beta = character(),
  max_se_metric   = numeric(), mover_se   = character(), metric_used_se   = character(),
  dLL = numeric(), dLL_rel = numeric(),
  stringsAsFactors = FALSE
)

prev_fit   <- NULL
prev_coefs <- NULL
Rcurr <- as.integer(R_start)

repeat{
  cat(sprintf("\n[gmnl] MIXL-E (corr=FALSE) con R = %d (Halton, BFGS) ...\n", Rcurr))
  t0 <- Sys.time()
  fitE <- fit_gmnl_safe(form_mixlE2, data = data, ranp = ranp,
                        R = Rcurr, corr = FALSE, start = prev_coefs)
  t1 <- Sys.time()
  llE <- len1(logLik(fitE))
  cat(sprintf("  Hecho en %.1f s | logLik = %.3f\n",
              as.numeric(difftime(t1, t0, units="secs")), llE))

  fits_E[[as.character(Rcurr)]] <- fitE
  coefs_now <- tryCatch(coef(fitE), error=function(e) NULL)

  if(is.null(prev_fit)){
    diag_tbl_E <- rbind(diag_tbl_E, data.frame(
      R = Rcurr, logLik = llE,
      max_beta_metric = NA_real_, mover_beta = NA_character_, metric_used_beta = NA_character_,
      max_se_metric   = NA_real_, mover_se   = NA_character_, metric_used_se   = NA_character_,
      dLL = NA_real_, dLL_rel = NA_real_
    ))
  } else {
    cmp <- compare_models_hybrid(fitE, prev_fit, beta_switch, se_switch)
    dLL_abs <- len1(cmp$dll)
    dLL_rel <- abs(dLL_abs) / (abs(len1(logLik(prev_fit))) + 1e-12)

    cat(sprintf("  Summing up: max β-metric=%.6f | max SE-metric=%.6f | ΔLL_rel=%.6e\n",
                len1(cmp$max_beta), len1(cmp$max_se), dLL_rel))

    row <- data.frame(
      R = Rcurr, logLik = llE,
      max_beta_metric = len1(cmp$max_beta),
      mover_beta      = cmp$top_beta$name,
      metric_used_beta= cmp$top_beta$used,
      max_se_metric   = len1(cmp$max_se),
      mover_se        = cmp$top_se$name,
      metric_used_se  = cmp$top_se$used,
      dLL             = dLL_abs,
      dLL_rel         = dLL_rel,
      stringsAsFactors = FALSE
    )

    # ---- MEAN VS. SD separate criteria ----
    b_new <- coef(fitE); b_old <- coef(prev_fit)
    alln  <- union(names(b_new), names(b_old))
    b_new <- b_new[alln]; b_old <- b_old[alln]
    is_sd <- grepl("^sd\\.", alln)

    db_abs <- abs(b_new - b_old)
    db_rel <- abs(b_new - b_old) / (abs(b_old) + 1e-12)

    # Mean
    db_abs_mean <- db_abs[!is_sd]; db_rel_mean <- db_rel[!is_sd]; b_old_mean <- b_old[!is_sd]
    beta_ok_mean <- all(ifelse(abs(b_old_mean) < beta_switch,
                               db_abs_mean < beta_abs_tol_mean,
                               db_rel_mean < beta_rel_tol_mean), na.rm = TRUE)

    # Deviations
    db_abs_sd <- db_abs[is_sd]; db_rel_sd <- db_rel[is_sd]; b_old_sd <- b_old[is_sd]
    beta_ok_sd <- all(ifelse(abs(b_old_sd) < beta_switch,
                             db_abs_sd < beta_abs_tol_sd,
                             db_rel_sd < beta_rel_tol_sd), na.rm = TRUE)

    beta_ok <- beta_ok_mean && beta_ok_sd
    se_ok   <- ifelse(row$metric_used_se == "rel",
                      row$max_se_metric < se_rel_tol,
                      row$max_se_metric < se_abs_tol)
    ll_ok   <- row$dLL_rel < ll_rel_tol

    diag_tbl_E <- rbind(diag_tbl_E, row)

    # ---- Stopping conditions ----
    if (!is.na(beta_ok) && !is.na(se_ok) && !is.na(ll_ok) && beta_ok && se_ok && ll_ok) {
      cat(">> The model reach stability (mean, sd, SE y ΔLL_rel). Stop.\n")
      prev_fit   <- fitE
      prev_coefs <- coefs_now
      break
    }
  }

  # ---- UPPER LIMIT CONTROL ----
  if (Rcurr + step_R > max_R){
    cat(sprintf(">> The model reach max_R=%d without reaching tolerance criteria. Stop.\n", max_R))
    prev_fit   <- fitE
    prev_coefs <- coefs_now
    break
  }

  prev_fit   <- fitE
  prev_coefs <- coefs_now
  Rcurr      <- Rcurr + step_R
}

# ===============================
# 5) Final model
# ===============================
R_final_E   <- diag_tbl_E$R[nrow(diag_tbl_E)]
mixlESP <- fits_E[[as.character(R_final_E)]]

cat(sprintf("\n[MIXL-E final)] R = %s | logLik = %.3f\n",
            R_final_E, len1(logLik(mixlESP))))
summary(mixlESP)

# Iteration's diagnose
diag_tbl_E

```
#### Reduced mixed logit 

```{r}
#set.seed(1685)semilla del modelo reportado, no borrar
set.seed(0611) #nueva semilla para comprobar estabilidad
# ===============================
# 1) SPECIFICATION: MIXL REDUCED ESP
# Project and Specie remain in utility, but are treated as fixed parameters
# Random parameters: Location and ASC_accept only
# ===============================

form_mixl_redESP <- Choice ~ 0 + Project + Specie + Location + Price +
  ASC_accept + ASC_Att + ASC_SN + ASC_PBC + ASC_PYME + ASC_Service

ranp_redESP <- c(
  Location   = "n",
  ASC_accept = "n"
)

data_redESP <- df_mld_ESP

# ===============================
# 2) LOOP PARAMETERS
# ===============================

R_start_redESP <- 3400
step_R_redESP  <- 200
max_R_redESP   <- 4000

# ---------------------------------------------------------------------------
# Tolerance thresholds — same criteria as the full Spain mixed logit loop.
# Full APA 6th references are listed separately in REFERENCES.md.
#
# beta_switch / se_switch = 0.05: use absolute differences near zero and
# relative differences otherwise (Train, 2009).
# Mean parameters = 0.001: stability in approximately the third significant
# digit across draw counts (Hensher & Greene, 2003; Train, 2009).
# SD parameters = 0.005: looser criterion for mixing-distribution standard
# deviations, which are more simulation-noisy (Czajkowski & Budzinski, 2019).
# SE tolerance = 0.010 absolute / 0.020 relative: evaluate simulation error
# relative to the SE magnitude itself (Hess, Train & Polak, 2006).
# Log-likelihood tolerance = 1e-5: negligible relative change across
# consecutive draw counts (Train, 2009; Czajkowski & Budzinski, 2019).
# ---------------------------------------------------------------------------

beta_switch_redESP        <- 0.05

beta_abs_tol_mean_redESP  <- 0.001
beta_rel_tol_mean_redESP  <- 0.001

beta_abs_tol_sd_redESP    <- 0.005
beta_rel_tol_sd_redESP    <- 0.005

se_switch_redESP          <- 0.05
se_abs_tol_redESP         <- 0.010
se_rel_tol_redESP         <- 0.020

ll_rel_tol_redESP         <- 1e-5

# ===============================
# 3) SAFE WRAPPER WITH BFGS
# ===============================

fit_gmnl_safe_redESP <- function(formula, data, ranp, R, corr = FALSE, start = NULL){
  gmnl(
    formula,
    data = data,
    model = "mixl",
    ranp = ranp,
    panel = TRUE,
    R = R,
    halton = list(prime = c(2, 3)),
    correlation = corr,
    method = "bfgs",
    start = start
  )
}

# ===============================
# 4) STABILISATION LOOP
# ===============================

fits_redESP <- list()

diag_tbl_R_redESP <- data.frame(
  R = integer(),
  logLik = numeric(),
  max_beta_metric = numeric(),
  mover_beta = character(),
  metric_used_beta = character(),
  max_se_metric = numeric(),
  mover_se = character(),
  metric_used_se = character(),
  dLL = numeric(),
  dLL_rel = numeric(),
  stringsAsFactors = FALSE
)

prev_fit_redESP   <- NULL
prev_coefs_redESP <- NULL
Rcurr_redESP      <- as.integer(R_start_redESP)

repeat{

  cat(sprintf(
    "\n[gmnl] MIXL-ESP reduced (corr=FALSE) con R = %d (Halton, BFGS) ...\n",
    Rcurr_redESP
  ))

  t0_redESP <- Sys.time()

  fit_redESP_iter <- fit_gmnl_safe_redESP(
    formula = form_mixl_redESP,
    data    = data_redESP,
    ranp    = ranp_redESP,
    R       = Rcurr_redESP,
    corr    = FALSE,
    start   = prev_coefs_redESP
  )

  t1_redESP <- Sys.time()

  ll_redESP_iter <- len1(logLik(fit_redESP_iter))

  cat(sprintf(
    "  Hecho en %.1f s | logLik = %.3f\n",
    as.numeric(difftime(t1_redESP, t0_redESP, units = "secs")),
    ll_redESP_iter
  ))

  fits_redESP[[as.character(Rcurr_redESP)]] <- fit_redESP_iter

  coefs_now_redESP <- tryCatch(
    coef(fit_redESP_iter),
    error = function(e) NULL
  )

  if(is.null(prev_fit_redESP)){

    diag_tbl_R_redESP <- rbind(
      diag_tbl_R_redESP,
      data.frame(
        R = Rcurr_redESP,
        logLik = ll_redESP_iter,
        max_beta_metric = NA_real_,
        mover_beta = NA_character_,
        metric_used_beta = NA_character_,
        max_se_metric = NA_real_,
        mover_se = NA_character_,
        metric_used_se = NA_character_,
        dLL = NA_real_,
        dLL_rel = NA_real_,
        stringsAsFactors = FALSE
      )
    )

  } else {

    cmp_redESP <- compare_models_hybrid(
      fit_redESP_iter,
      prev_fit_redESP,
      beta_switch_redESP,
      se_switch_redESP
    )

    dLL_abs_redESP <- len1(cmp_redESP$dll)

    dLL_rel_redESP <- abs(dLL_abs_redESP) /
      (abs(len1(logLik(prev_fit_redESP))) + 1e-12)

    cat(sprintf(
      "  Summing up: max β-metric=%.6f | max SE-metric=%.6f | ΔLL_rel=%.6e\n",
      len1(cmp_redESP$max_beta),
      len1(cmp_redESP$max_se),
      dLL_rel_redESP
    ))

    row_redESP <- data.frame(
      R = Rcurr_redESP,
      logLik = ll_redESP_iter,
      max_beta_metric = len1(cmp_redESP$max_beta),
      mover_beta = cmp_redESP$top_beta$name,
      metric_used_beta = cmp_redESP$top_beta$used,
      max_se_metric = len1(cmp_redESP$max_se),
      mover_se = cmp_redESP$top_se$name,
      metric_used_se = cmp_redESP$top_se$used,
      dLL = dLL_abs_redESP,
      dLL_rel = dLL_rel_redESP,
      stringsAsFactors = FALSE
    )

    # ---- MEAN VS. SD separate criteria ----

    b_new_redESP <- coef(fit_redESP_iter)
    b_old_redESP <- coef(prev_fit_redESP)

    alln_redESP <- union(names(b_new_redESP), names(b_old_redESP))

    b_new_redESP <- b_new_redESP[alln_redESP]
    b_old_redESP <- b_old_redESP[alln_redESP]

    is_sd_redESP <- grepl("^sd\\.", alln_redESP)

    db_abs_redESP <- abs(b_new_redESP - b_old_redESP)

    db_rel_redESP <- abs(b_new_redESP - b_old_redESP) /
      (abs(b_old_redESP) + 1e-12)

    # Mean parameters

    db_abs_mean_redESP <- db_abs_redESP[!is_sd_redESP]
    db_rel_mean_redESP <- db_rel_redESP[!is_sd_redESP]
    b_old_mean_redESP  <- b_old_redESP[!is_sd_redESP]

    beta_ok_mean_redESP <- all(
      ifelse(
        abs(b_old_mean_redESP) < beta_switch_redESP,
        db_abs_mean_redESP < beta_abs_tol_mean_redESP,
        db_rel_mean_redESP < beta_rel_tol_mean_redESP
      ),
      na.rm = TRUE
    )

    # Standard deviation parameters

    db_abs_sd_redESP <- db_abs_redESP[is_sd_redESP]
    db_rel_sd_redESP <- db_rel_redESP[is_sd_redESP]
    b_old_sd_redESP  <- b_old_redESP[is_sd_redESP]

    beta_ok_sd_redESP <- all(
      ifelse(
        abs(b_old_sd_redESP) < beta_switch_redESP,
        db_abs_sd_redESP < beta_abs_tol_sd_redESP,
        db_rel_sd_redESP < beta_rel_tol_sd_redESP
      ),
      na.rm = TRUE
    )

    beta_ok_redESP <- beta_ok_mean_redESP && beta_ok_sd_redESP

    se_ok_redESP <- ifelse(
      row_redESP$metric_used_se == "rel",
      row_redESP$max_se_metric < se_rel_tol_redESP,
      row_redESP$max_se_metric < se_abs_tol_redESP
    )

    ll_ok_redESP <- row_redESP$dLL_rel < ll_rel_tol_redESP

    diag_tbl_R_redESP <- rbind(diag_tbl_R_redESP, row_redESP)

    # ---- Stopping conditions ----

    if (
      !is.na(beta_ok_redESP) &&
      !is.na(se_ok_redESP) &&
      !is.na(ll_ok_redESP) &&
      beta_ok_redESP &&
      se_ok_redESP &&
      ll_ok_redESP
    ) {
      cat(">> The reduced MIXL-ESP model reached stability (mean, sd, SE and ΔLL_rel). Stop.\n")

      prev_fit_redESP   <- fit_redESP_iter
      prev_coefs_redESP <- coefs_now_redESP

      break
    }
  }

  # ---- UPPER LIMIT CONTROL ----

  if (Rcurr_redESP + step_R_redESP > max_R_redESP){

    cat(sprintf(
      ">> The reduced MIXL-ESP model reached max_R=%d without reaching all tolerance criteria. Stop.\n",
      max_R_redESP
    ))

    prev_fit_redESP   <- fit_redESP_iter
    prev_coefs_redESP <- coefs_now_redESP

    break
  }

  prev_fit_redESP   <- fit_redESP_iter
  prev_coefs_redESP <- coefs_now_redESP
  Rcurr_redESP      <- Rcurr_redESP + step_R_redESP
}

# ===============================
# 5) FINAL MODEL: REDUCED MIXL ESP
# ===============================

R_final_redESP <- diag_tbl_R_redESP$R[nrow(diag_tbl_R_redESP)]

mixlESP_red <- fits_redESP[[as.character(R_final_redESP)]]

cat(sprintf(
  "\n[MIXL-ESP reduced final] R = %s | logLik = %.3f\n",
  R_final_redESP,
  len1(logLik(mixlESP_red))
))

summary(mixlESP_red)

# Iteration diagnosis
diag_tbl_R_redESP
```


### LR test to choose random parameters
#### Registry
```{r}
# ===============================
# 6) MODEL COMPARISON
# ===============================

ll_full_RHC <- as.numeric(logLik(mixlRHC))
ll_red_RHC  <- as.numeric(logLik(mixlRHC_red))

LR_red_vs_full_RHC <- 2 * (ll_full_RHC - ll_red_RHC)

df_red_vs_full_RHC <- length(coef(mixlRHC)) - length(coef(mixlRHC_red))

p_LRT_standard_RHC <- pchisq(
  LR_red_vs_full_RHC,
  df = df_red_vs_full_RHC,
  lower.tail = FALSE
)

comparison_redRHC <- data.frame(
  model = c("Full MIXL", "Reduced MIXL"),
  logLik = c(ll_full_RHC, ll_red_RHC),
  npar = c(length(coef(mixlRHC)), length(coef(mixlRHC_red))),
  AIC = c(AIC(mixlRHC), AIC(mixlRHC_red)),
  BIC = c(BIC(mixlRHC), BIC(mixlRHC_red))
)

comparison_redRHC

cat("\nLikelihood-ratio test: full vs reduced MIXL\n")
cat("LR statistic:", LR_red_vs_full_RHC, "\n")
cat("df:", df_red_vs_full_RHC, "\n")
cat("standard p-value:", p_LRT_standard_RHC, "\n")


# ===============================
# 7) COEFFICIENT STABILITY: FULL VS REDUCED
# ===============================

coef_full_RHC <- coef(mixlRHC)
coef_red_RHC  <- coef(mixlRHC_red)

common_coef_RHC <- intersect(names(coef_full_RHC), names(coef_red_RHC))

coef_compare_redRHC <- data.frame(
  parameter = common_coef_RHC,
  full = coef_full_RHC[common_coef_RHC],
  reduced = coef_red_RHC[common_coef_RHC],
  abs_diff = abs(coef_red_RHC[common_coef_RHC] - coef_full_RHC[common_coef_RHC]),
  rel_diff = abs(coef_red_RHC[common_coef_RHC] - coef_full_RHC[common_coef_RHC]) /
    (abs(coef_full_RHC[common_coef_RHC]) + 1e-12)
)

coef_compare_redRHC


```

#### Spain

```{r}
# ===============================
# 6) MODEL COMPARISON: FULL VS REDUCED ESP
# ===============================

ll_full_ESP <- as.numeric(logLik(mixlESP))
ll_red_ESP  <- as.numeric(logLik(mixlESP_red))

LR_red_vs_full_ESP <- 2 * (ll_full_ESP - ll_red_ESP)

df_red_vs_full_ESP <- length(coef(mixlESP)) - length(coef(mixlESP_red))

p_LRT_standard_ESP <- pchisq(
  LR_red_vs_full_ESP,
  df = df_red_vs_full_ESP,
  lower.tail = FALSE
)

comparison_redESP <- data.frame(
  model = c("Full MIXL ESP", "Reduced MIXL ESP"),
  logLik = c(ll_full_ESP, ll_red_ESP),
  npar = c(length(coef(mixlESP)), length(coef(mixlESP_red))),
  AIC = c(AIC(mixlESP), AIC(mixlESP_red)),
  BIC = c(BIC(mixlESP), BIC(mixlESP_red))
)

comparison_redESP

cat("\nLikelihood-ratio test: full vs reduced MIXL ESP\n")
cat("LR statistic:", LR_red_vs_full_ESP, "\n")
cat("df:", df_red_vs_full_ESP, "\n")
cat("standard p-value:", p_LRT_standard_ESP, "\n")

# ===============================
# 7) COEFFICIENT STABILITY: FULL VS REDUCED ESP
# ===============================

coef_full_ESP <- coef(mixlESP)
coef_red_ESP  <- coef(mixlESP_red)

common_coef_ESP <- intersect(names(coef_full_ESP), names(coef_red_ESP))

coef_compare_redESP <- data.frame(
  parameter = common_coef_ESP,
  full = coef_full_ESP[common_coef_ESP],
  reduced = coef_red_ESP[common_coef_ESP],
  abs_diff = abs(coef_red_ESP[common_coef_ESP] - coef_full_ESP[common_coef_ESP]),
  rel_diff = abs(coef_red_ESP[common_coef_ESP] - coef_full_ESP[common_coef_ESP]) /
    (abs(coef_full_ESP[common_coef_ESP]) + 1e-12)
)

coef_compare_redESP

```



### Model fit
```{r}
AIC(mixlRHC_red)
AIC(mixlESP_red)
BIC(mixlRHC_red)
BIC(mixlESP_red)

##Mixed logit
# Log-likelihood 
ll_modelA <- as.numeric(logLik(mixlRHC_red))

# Log-likelihood null model 
ll_nullA <- as.numeric(logLik(mlogit(Choice ~ 1, data = df_mld_RHC)))

# McFadden's pseudo-R2 for discrete choice model fit (McFadden, 1974).
# Full reference: REFERENCES.md.
R2_McFaddenA <- 1 - (ll_modelA / ll_nullA)
R2_McFaddenA


#Segment B
# Log-likelihood del modelo estimado
ll_modelB <- as.numeric(logLik(mixlESP_red))

# Log-likelihood null model
ll_nullB <- as.numeric(logLik(mlogit(Choice ~ 1, data = df_mld_ESP)))

# McFadden's pseudo-R2 for discrete choice model fit (McFadden, 1974).
# Full reference: REFERENCES.md.
R2_McFaddenB <- 1 - (ll_modelB / ll_nullB)
R2_McFaddenB

```

## Willigness to pay estimations

### Krinsky-Robb

#### Registry

```{r}
##WTP
# Krinsky-Robb WTP intervals: simulate parameter draws from the estimated
# coefficient vector and variance-covariance matrix (Krinsky & Robb, 1986,
# 1990). Full references: REFERENCES.md.
# Coefs y VCOV


# WTP attributes
attrs <- c("Project","Specie","Location")


set.seed(0611)
SR <- 20000  # nº de draws
bR   <- coef(mixlRHC_red)
VR   <- vcov(mixlRHC_red)

# Takes draws from the parametric vector
draws <- MASS::mvrnorm(n = SR, mu = bR, Sigma = VR)

# Helper to extraxt values in named columns
col_ix <- function(nm) match(nm, names(bR))

wtp_kr <- lapply(attrs, function(attr){
  bk <- draws[, col_ix(attr)]
  bp <- draws[, col_ix("Price")]
  w  <- -2* bk / bp
  data.frame(attribute=attr,
             WTP_mean = mean(w, na.rm=TRUE),
             Q2.5 = quantile(w, 0.025, na.rm=TRUE),
             Q97.5 = quantile(w, 0.975, na.rm=TRUE))
})
wtp_table_kr <- bind_rows(wtp_kr)
print(wtp_table_kr)

bp <- draws[, col_ix("Price")]

wtp_total_draws <- rowSums(
  sapply(attrs, function(attr) {
    bk <- draws[, col_ix(attr)]
    -2 * bk / bp
  })
)

wtp_total_row <- data.frame(
  attribute = "Total",
  WTP_mean  = mean(wtp_total_draws, na.rm = TRUE),
  Q2.5      = quantile(wtp_total_draws, 0.025, na.rm = TRUE),
  Q97.5     = quantile(wtp_total_draws, 0.975, na.rm = TRUE)
)

wtp_table_kr <- bind_rows(wtp_table_kr, wtp_total_row)
print(wtp_table_kr)

```

#### Spain

```{r}
# ===============================
# WTP ESPAÑA — mixlESP_red2
# ===============================
# Krinsky-Robb WTP intervals: simulate parameter draws from the estimated
# coefficient vector and variance-covariance matrix (Krinsky & Robb, 1986,
# 1990). Full references: REFERENCES.md.
attrsE <- c("Project", "Specie", "Location")
set.seed(0611)
SE <- 20000

bE <- coef(mixlESP_red)
VE <- vcov(mixlESP_red)

# col_ix PROPIO para España — apunta a bE, no a bR
col_ix_E <- function(nm) match(nm, names(bE))

drawsE <- MASS::mvrnorm(n = SE, mu = bE, Sigma = VE)

# WTP por atributo
wtp_kr_E <- lapply(attrsE, function(attrE) {
  bkE <- drawsE[, col_ix_E(attrE)]
  bpE <- drawsE[, col_ix_E("Price")]
  wE  <- -2 * bkE / bpE          # ajusta a -2*bkE/bpE si tu Price está dividido por 2
  data.frame(
    attribute = attrE,
    WTP_mean  = mean(wE, na.rm = TRUE),
    Q2.5      = quantile(wE, 0.025, na.rm = TRUE),
    Q97.5     = quantile(wE, 0.975, na.rm = TRUE)
  )
})

# WTP total (draw a draw)
bpE_vec <- drawsE[, col_ix_E("Price")]
wtp_total_draws_E <- rowSums(
  sapply(attrsE, function(attrE) {
    bkE <- drawsE[, col_ix_E(attrE)]
    -2 * bkE / bpE_vec             # mismo factor que arriba
  })
)

wtp_total_row_E <- data.frame(
  attribute = "Total",
  WTP_mean  = mean(wtp_total_draws_E, na.rm = TRUE),
  Q2.5      = quantile(wtp_total_draws_E, 0.025, na.rm = TRUE),
  Q97.5     = quantile(wtp_total_draws_E, 0.975, na.rm = TRUE)
)

wtp_table_kr_E <- bind_rows(wtp_kr_E, wtp_total_row_E)
print(wtp_table_kr_E)


```

## Willigness to pay estimations: 10/100€ tCO2

### Krinsky-Robb

#### Registry

```{r}
##WTP
# Coefs y VCOV


# WTP attributes
attrs <- c("Project","Specie","Location")


set.seed(0611)
SR <- 20000  # nº de draws
bR   <- coef(mixlRHC_red)
VR   <- vcov(mixlRHC_red)

# Takes draws from the parametric vector
draws <- MASS::mvrnorm(n = SR, mu = bR, Sigma = VR)

# Helper to extraxt values in named columns
col_ix <- function(nm) match(nm, names(bR))

wtp_kr <- lapply(attrs, function(attr){
  bk <- draws[, col_ix(attr)]
  bp <- draws[, col_ix("Price")]
  w  <- -2* bk / bp
  data.frame(attribute=attr,
             WTP_mean = mean(w, na.rm=TRUE),
             Q2.5 = quantile(w, 0.025, na.rm=TRUE),
             Q97.5 = quantile(w, 0.975, na.rm=TRUE))
})
wtp_table_kr <- bind_rows(wtp_kr)
print(wtp_table_kr)

bp <- draws[, col_ix("Price")]

wtp_total_draws <- rowSums(
  sapply(attrs, function(attr) {
    bk <- draws[, col_ix(attr)]
    -2 * bk / bp
  })
)

wtp_total_row <- data.frame(
  attribute = "Total",
  WTP_mean  = mean(wtp_total_draws, na.rm = TRUE),
  Q2.5      = quantile(wtp_total_draws, 0.025, na.rm = TRUE),
  Q97.5     = quantile(wtp_total_draws, 0.975, na.rm = TRUE)
)

wtp_table_kr <- bind_rows(wtp_table_kr, wtp_total_row)
print(wtp_table_kr)

```

#### Spain

```{r}
# ===============================
# WTP ESPAÑA — mixlESP_red2
# ===============================
attrsE <- c("Project", "Specie", "Location")
set.seed(0611)
SE <- 20000

bE <- coef(mixlESP_red)
VE <- vcov(mixlESP_red)

# col_ix PROPIO para España — apunta a bE, no a bR
col_ix_E <- function(nm) match(nm, names(bE))

drawsE <- MASS::mvrnorm(n = SE, mu = bE, Sigma = VE)

# WTP por atributo
wtp_kr_E <- lapply(attrsE, function(attrE) {
  bkE <- drawsE[, col_ix_E(attrE)]
  bpE <- drawsE[, col_ix_E("Price")]
  wE  <- -2 * bkE / bpE          # ajusta a -2*bkE/bpE si tu Price está dividido por 2
  data.frame(
    attribute = attrE,
    WTP_mean  = mean(wE, na.rm = TRUE),
    Q2.5      = quantile(wE, 0.025, na.rm = TRUE),
    Q97.5     = quantile(wE, 0.975, na.rm = TRUE)
  )
})

# WTP total (draw a draw)
bpE_vec <- drawsE[, col_ix_E("Price")]
wtp_total_draws_E <- rowSums(
  sapply(attrsE, function(attrE) {
    bkE <- drawsE[, col_ix_E(attrE)]
    -2 * bkE / bpE_vec             # mismo factor que arriba
  })
)

wtp_total_row_E <- data.frame(
  attribute = "Total",
  WTP_mean  = mean(wtp_total_draws_E, na.rm = TRUE),
  Q2.5      = quantile(wtp_total_draws_E, 0.025, na.rm = TRUE),
  Q97.5     = quantile(wtp_total_draws_E, 0.975, na.rm = TRUE)
)

wtp_table_kr_E <- bind_rows(wtp_kr_E, wtp_total_row_E)
print(wtp_table_kr_E)


```





## Latent class model: Number of classes selection

### Data importation
```{r}
database <- as.data.frame(
  read_excel(data_path,
             sheet = "Todo_II") %>%
    mutate(
      Project    = ifelse(Optout == 1 & Project  == -4, 0, Project),
      Specie     = ifelse(Optout == 1 & Specie   == -4, 0, Specie),
      Location   = ifelse(Optout == 1 & Location == -4, 0, Location),
      Price      = ifelse(Optout == 1 & Price    == -4, 0, Price),
      ASC_accept = as.numeric(ifelse(Optout == 1, 0, 1)),
      alt_id     = ifelse(Optout == 0, 1, 2),
      chid       = paste(Id, ParentQuestionId, QuestionId, sep = "_")
    ) %>%
    group_by(chid) %>%
    mutate(choiceVar = alt_id[which(Choice == 1)][1]) %>%
    ungroup()
)

n_indiv <- length(unique(database$Id))
n_cs    <- length(unique(database$chid))
n_obs   <- nrow(database)
cat("Individuos:", n_indiv, "| Choice sets:", n_cs,
    "| Observaciones:", n_obs, "\n")


# LL(0) correct: -n_obs * log(J), J = 2 alternatives
LL0_correct <- -n_obs * log(2)

```


### Two classes model

#### Diagnosis function


```{r}
diagnosticos_LC <- function(model, C, apollo_probabilities,
                             apollo_inputs, n_indiv) {

  # Latent class selection diagnostics combine information criteria
  # (Akaike, 1974; Schwarz, 1978), McFadden rho2 (McFadden, 1974),
  # posterior classification diagnostics and entropy, plus convergence and
  # a valid covariance matrix. APCP/AvePP and OCC follow Nagin (2005) and
  # PROC LCA guidance (Lanza et al., 2007, 2015); entropy follows Celeux &
  # Soromenho (1996), Nylund, Asparouhov & Muthen (2007), and Latent GOLD
  # reference guidance (Vermunt & Magidson, 2005).
  # Full references: REFERENCES.md.

  ll  <- as.numeric(model$LLout[1])
  k   <- length(model$estimate)
  nms <- names(model$LLout)

  # LL0 correct (n_obs, no n_cs)
  LL0 <- LL0_correct

  # LLc from Apollo
  LLc <- if ("LL(c)" %in% nms) as.numeric(model$LLout["LL(c)"])
         else if ("LLc" %in% nms) as.numeric(model$LLout["LLc"])
         else ll

  aic    <- -2 * ll + 2 * k
  bic    <- -2 * ll + k * log(n_cs)
  rho2_0 <- as.numeric(1 - ll / LL0)
  rho2_c <- as.numeric(1 - ll / LLc)

  unconditionals <- apollo_lcUnconditionals(
    model, apollo_probabilities, apollo_inputs)
  n_cl        <- length(unconditionals[["pi_values"]])
  class_sizes <- colMeans(matrix(unlist(unconditionals[["pi_values"]]),
                                  ncol = n_cl, byrow = FALSE))

  conditionals <- apollo_lcConditionals(
    model, apollo_probabilities, apollo_inputs)

  avg_post_prob <- colMeans(conditionals[, 2:(C + 1)])

  modal_assign <- apply(conditionals[, 2:(C + 1)], 1, which.max)
  modal_counts <- as.vector(table(factor(modal_assign, levels = 1:C)))
  modal_shares <- modal_counts / n_indiv

  post_mat  <- as.matrix(conditionals[, 2:(C + 1)])
  post_mat  <- pmax(post_mat, 1e-10)
  entropy_n <- -rowSums(post_mat * log(post_mat))
  entropy   <- 1 - mean(entropy_n) / log(C)

  apcp_modal <- numeric(C)
  for (q in seq_len(C)) {
    idx <- which(modal_assign == q)
    apcp_modal[q] <- if (length(idx) > 0) mean(post_mat[idx, q]) else NA
  }
  apcp_modal <- pmin(apcp_modal, 0.9999)
  occ <- (apcp_modal / (1 - apcp_modal)) / (class_sizes / (1 - class_sizes))

  list(C = C, LL = round(ll, 2), AIC = round(aic, 2),
       BIC = round(bic, 2), npar = k,
       rho2_0 = round(rho2_0, 4), rho2_c = round(rho2_c, 4),
       entropy  = round(entropy, 4),
       class_sizes   = round(class_sizes,   4),
       avg_post_prob = round(avg_post_prob, 4),  # APCP global (para tabla)
       apcp_modal    = round(apcp_modal,    4),  # APCP modal  (para OCC)
       modal_shares  = round(modal_shares,  4),
       occ           = round(occ, 3))
}

```

#### Apollo: beta, fixed, lcPars


```{r}
cat("\n", strrep("=", 60), "\n")
cat("  ESTIMATING C = 2 CLASSES\n")
cat(strrep("=", 60), "\n")

apollo_control <- list(
  modelName = "LC_2cl", modelDescr = "LC 2 clasSes",
  indivID = "Id", mixing = FALSE, nCores = 1,
  outputDirectory = "output"
)

apollo_beta <- c(
  asc_a = 1.0, asc_b = -1.0,
  b_Project_a  = 0.10, b_Specie_a  = 0.10,
  b_Location_a = 0.20, b_Price_log_a = -3.91,
  b_Project_b  = 0.10, b_Specie_b  = 0.10,
  b_Location_b = 0.20, b_Price_log_b = -3.91,
  delta_a = 0.0, delta_b = 0.0
)
apollo_fixed <- "delta_b"

apollo_lcPars <- function(apollo_beta, apollo_inputs) {
  lcpars <- list(
    asc        = list(asc_a,        asc_b),
    b_Project  = list(b_Project_a,  b_Project_b),
    b_Specie   = list(b_Specie_a,   b_Specie_b),
    b_Location = list(b_Location_a, b_Location_b),
    b_Price    = list(-exp(b_Price_log_a), -exp(b_Price_log_b))
  )
  V_class <- list(class_a = delta_a, class_b = delta_b)
  lcpars[["pi_values"]] <- apollo_classAlloc(list(
    classes = c(class_a = 1, class_b = 2),
    utilities = V_class, componentName = "classAlloc"))
  return(lcpars)
}
```

#### Apollo probabilities and inputs
```{r}
apollo_probabilities <- function(apollo_beta, apollo_inputs,
                                 functionality = "estimate") {
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  P     <- list()
  mnl_s <- list(alternatives = c(project = 1, optout = 2),
                avail        = list(project = 1, optout = 1),
                choiceVar    = choiceVar)
  for (s in 1:2) {
    V <- list()
    V[["project"]] <- asc[[s]] * ASC_accept +
      b_Project[[s]] * Project + b_Specie[[s]] * Specie +
      b_Location[[s]] * Location + b_Price[[s]] * Price
    V[["optout"]] <- 0
    mnl_s$utilities     <- V
    mnl_s$componentName <- paste0("Class_", s)
    P[[paste0("Class_", s)]] <- apollo_mnl(mnl_s, functionality)
    P[[paste0("Class_", s)]] <- apollo_panelProd(
      P[[paste0("Class_", s)]], apollo_inputs, functionality)
  }
  P[["model"]] <- apollo_lc(
    list(inClassProb = P, classProb = pi_values),
    apollo_inputs, functionality)
  return(apollo_prepareProb(P, apollo_inputs, functionality))
}

apollo_inputs <- apollo_validateInputs()
model_2cl <- apollo_estimate(
  apollo_beta, apollo_fixed,
  apollo_probabilities, apollo_inputs,
  estimate_settings = list(estimationRoutine = "bgw",
    maxIterations = 2000, writeIter = FALSE, numDeriv = TRUE))
apollo_saveOutput(model_2cl)
apollo_sink()
```



#### 2 classes results

```{r}
res_2 <- diagnosticos_LC(model_2cl, 2, apollo_probabilities,
                          apollo_inputs, n_indiv)
```




### Three classes model
#### Apollo: beta, fixed, lcPars


```{r}
cat("\n", strrep("=", 60), "\n")
cat("  ESTIMANDO C = 3 CLASSES\n")
cat(strrep("=", 60), "\n")

apollo_control <- list(
  modelName = "LC_3cl", modelDescr = "LC 3 classes",
  indivID = "Id", mixing = FALSE, nCores = 1,
  outputDirectory = "output"
)

apollo_beta <- c(
  asc_a = 1.0, asc_b = -1.0, asc_c = 0.2,
  b_Project_a  = 0.10, b_Specie_a  = 0.10,
  b_Location_a = 0.20, b_Price_log_a = -3.91,
  b_Project_b  = 0.10, b_Specie_b  = 0.10,
  b_Location_b = 0.20, b_Price_log_b = -3.91,
  b_Project_c  = 0.10, b_Specie_c  = 0.10,
  b_Location_c = 0.20, b_Price_log_c = -3.91,
  delta_a = 0.0, delta_b = 0.0, delta_c = 0.0
)
apollo_fixed <- "delta_c"

apollo_lcPars <- function(apollo_beta, apollo_inputs) {
  lcpars <- list(
    asc        = list(asc_a,        asc_b,        asc_c),
    b_Project  = list(b_Project_a,  b_Project_b,  b_Project_c),
    b_Specie   = list(b_Specie_a,   b_Specie_b,   b_Specie_c),
    b_Location = list(b_Location_a, b_Location_b, b_Location_c),
    b_Price    = list(-exp(b_Price_log_a),
                      -exp(b_Price_log_b),
                      -exp(b_Price_log_c))
  )
  V_class <- list(class_a = delta_a,
                  class_b = delta_b,
                  class_c = delta_c)
  lcpars[["pi_values"]] <- apollo_classAlloc(list(
    classes = c(class_a = 1, class_b = 2, class_c = 3),
    utilities = V_class, componentName = "classAlloc"))
  return(lcpars)
}

```


#### Apollo: probabilities and inputs

```{r}

apollo_probabilities <- function(apollo_beta, apollo_inputs,
                                 functionality = "estimate") {
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  P     <- list()
  mnl_s <- list(alternatives = c(project = 1, optout = 2),
                avail        = list(project = 1, optout = 1),
                choiceVar    = choiceVar)
  for (s in 1:3) {
    V <- list()
    V[["project"]] <- asc[[s]] * ASC_accept +
      b_Project[[s]] * Project + b_Specie[[s]] * Specie +
      b_Location[[s]] * Location + b_Price[[s]] * Price
    V[["optout"]] <- 0
    mnl_s$utilities     <- V
    mnl_s$componentName <- paste0("Class_", s)
    P[[paste0("Class_", s)]] <- apollo_mnl(mnl_s, functionality)
    P[[paste0("Class_", s)]] <- apollo_panelProd(
      P[[paste0("Class_", s)]], apollo_inputs, functionality)
  }
  P[["model"]] <- apollo_lc(
    list(inClassProb = P, classProb = pi_values),
    apollo_inputs, functionality)
  return(apollo_prepareProb(P, apollo_inputs, functionality))
}

apollo_inputs <- apollo_validateInputs()
model_3cl <- apollo_estimate(
  apollo_beta, apollo_fixed,
  apollo_probabilities, apollo_inputs,
  estimate_settings = list(estimationRoutine = "bgw",
    maxIterations = 2000, writeIter = FALSE, numDeriv = TRUE))
apollo_saveOutput(model_3cl)
apollo_sink()
```

#### 3 classes result
```{r}

res_3 <- diagnosticos_LC(model_3cl, 3, apollo_probabilities,
                          apollo_inputs, n_indiv)
```



### Four classes model

#### Apollo: beta, fixed, lcPars

```{r}

cat("\n", strrep("=", 60), "\n")
cat("  ESTIMANDO C = 4 CLASSES\n")
cat(strrep("=", 60), "\n")

apollo_control <- list(
  modelName = "LC_4cl", modelDescr = "LC 4 classes",
  indivID = "Id", mixing = FALSE, nCores = 1,
  outputDirectory = "output"
)

apollo_beta <- c(
  asc_a = 1.0, asc_b = -1.0, asc_c = 0.2, asc_d = -0.2,
  b_Project_a  = 0.10, b_Specie_a  = 0.10,
  b_Location_a = 0.20, b_Price_log_a = -3.91,
  b_Project_b  = 0.10, b_Specie_b  = 0.10,
  b_Location_b = 0.20, b_Price_log_b = -3.91,
  b_Project_c  = 0.10, b_Specie_c  = 0.10,
  b_Location_c = 0.20, b_Price_log_c = -3.91,
  b_Project_d  = 0.10, b_Specie_d  = 0.10,
  b_Location_d = 0.20, b_Price_log_d = -3.91,
  delta_a = 0.0, delta_b = 0.0,
  delta_c = 0.0, delta_d = 0.0
)
apollo_fixed <- "delta_d"

apollo_lcPars <- function(apollo_beta, apollo_inputs) {
  lcpars <- list(
    asc        = list(asc_a,        asc_b,
                      asc_c,        asc_d),
    b_Project  = list(b_Project_a,  b_Project_b,
                      b_Project_c,  b_Project_d),
    b_Specie   = list(b_Specie_a,   b_Specie_b,
                      b_Specie_c,   b_Specie_d),
    b_Location = list(b_Location_a, b_Location_b,
                      b_Location_c, b_Location_d),
    b_Price    = list(-exp(b_Price_log_a), -exp(b_Price_log_b),
                      -exp(b_Price_log_c), -exp(b_Price_log_d))
  )
  V_class <- list(class_a = delta_a, class_b = delta_b,
                  class_c = delta_c, class_d = delta_d)
  lcpars[["pi_values"]] <- apollo_classAlloc(list(
    classes = c(class_a = 1, class_b = 2,
                class_c = 3, class_d = 4),
    utilities = V_class, componentName = "classAlloc"))
  return(lcpars)
}
```


#### Apollo: probabilities and inputs

```{r}

apollo_probabilities <- function(apollo_beta, apollo_inputs,
                                 functionality = "estimate") {
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))
  P     <- list()
  mnl_s <- list(alternatives = c(project = 1, optout = 2),
                avail        = list(project = 1, optout = 1),
                choiceVar    = choiceVar)
  for (s in 1:4) {
    V <- list()
    V[["project"]] <- asc[[s]] * ASC_accept +
      b_Project[[s]] * Project + b_Specie[[s]] * Specie +
      b_Location[[s]] * Location + b_Price[[s]] * Price
    V[["optout"]] <- 0
    mnl_s$utilities     <- V
    mnl_s$componentName <- paste0("Class_", s)
    P[[paste0("Class_", s)]] <- apollo_mnl(mnl_s, functionality)
    P[[paste0("Class_", s)]] <- apollo_panelProd(
      P[[paste0("Class_", s)]], apollo_inputs, functionality)
  }
  P[["model"]] <- apollo_lc(
    list(inClassProb = P, classProb = pi_values),
    apollo_inputs, functionality)
  return(apollo_prepareProb(P, apollo_inputs, functionality))
}

apollo_inputs <- apollo_validateInputs()
model_4cl <- apollo_estimate(
  apollo_beta, apollo_fixed,
  apollo_probabilities, apollo_inputs,
  estimate_settings = list(estimationRoutine = "bgw",
    maxIterations = 2000, writeIter = FALSE, numDeriv = TRUE))
apollo_saveOutput(model_4cl)
apollo_sink()
```

#### 4 classes result
```{r}
res_4 <- diagnosticos_LC(model_4cl, 4, apollo_probabilities,
                        apollo_inputs, n_indiv)
```

### Class selection test
```{r}


cat("\n===    LLout Models' names ===\n")
cat("C=2:", names(model_2cl$LLout), "\n")
cat("C=3:", names(model_3cl$LLout), "\n")
cat("C=4:", names(model_4cl$LLout), "\n")

get_ll_metrics <- function(model) {
  ll  <- as.numeric(model$LLout[1])
  k   <- length(model$estimate)
  nms <- names(model$LLout)

  cat("  LLout:", paste(nms, round(model$LLout, 2),
                        sep = "=", collapse = " | "), "\n")

  LLc <- if ("LL(c)" %in% nms) as.numeric(model$LLout["LL(c)"])
         else if ("LLc" %in% nms) as.numeric(model$LLout["LLc"])
         else ll

  aic    <- -2 * ll + 2 * k
  bic    <- -2 * ll + k * log(n_cs)
  rho2_0 <- 1 - ll / LL0_correcto
  rho2_c <- 1 - ll / LLc

  cat(sprintf("  LL=%.2f | LL0=%.2f | LLc=%.2f | rho2_0=%.4f | rho2_c=%.4f\n",
              ll, LL0_correcto, LLc, rho2_0, rho2_c))

  list(ll = ll, k = k, aic = aic, bic = bic,
       rho2_0 = rho2_0, rho2_c = rho2_c)
}

cat("\n=== Criteria verification ===\n")
cat("C=2:\n"); m2 <- get_ll_metrics(model_2cl)
cat("C=3:\n"); m3 <- get_ll_metrics(model_3cl)
cat("C=4:\n"); m4 <- get_ll_metrics(model_4cl)

```


#### Convergence, test and parameters

```{r}

# ================================================================= #
#### TABLA FINAL                                                 ####
# ================================================================= #

results   <- list(res_2, res_3, res_4)
convergence <- c(TRUE, FALSE, FALSE)  # C=3 singular; C=4 sin Hessiana
bic_valid_min <- min(c(m2$bic, m3$bic, m4$bic)[convergence])
fit_table <- tibble(
  Classes   = c(2,    3,    4),
  LL        = c(m2$ll,    m3$ll,    m4$ll),
  npar      = c(m2$k,     m3$k,     m4$k),
  AIC       = c(m2$aic,   m3$aic,   m4$aic),
  BIC       = c(m2$bic,   m3$bic,   m4$bic),
  Rho2_0    = c(m2$rho2_0, m3$rho2_0, m4$rho2_0),
  Rho2_c    = c(m2$rho2_c, m3$rho2_c, m4$rho2_c),
  Entropy   = c(res_2$entropy, res_3$entropy, res_4$entropy),
  Converged = convergence
) %>%
  mutate(
    delta_AIC  = ifelse(Converged, round(AIC - min(AIC[Converged]), 2), NA),
    delta_BIC  = ifelse(Converged, round(BIC - bic_valid_min, 2), NA),
    LR_vs_prev = c(NA, round(2 * diff(LL), 2)),
    df_prev    = c(NA, diff(npar)),
    p_LR       = c(NA, mapply(function(lr, df) {
      if (is.na(lr) || is.na(df)) return(NA)
      round(pchisq(lr, df = df, lower.tail = FALSE), 4)
    }, LR_vs_prev[-1], df_prev[-1]))
  )

cat("\n", strrep("=", 70), "\n")
cat("  Final table\n")
cat(strrep("=", 70), "\n\n")
print(as.data.frame(fit_table), digits = 4)

cat("\n", strrep("=", 70), "\n")
cat("  Classification diagnostics\n")
cat(strrep("=", 70), "\n")

for (res in results) {
  C <- res$C
  cat(sprintf("\nModel with C = %d classes\n", C))
  
  for (j in seq_len(C)) {
    cat(sprintf(
      "  Clase %s: pi=%.3f | APCP_global=%.3f | APCP_modal=%.3f | OCC=%.2f\n",
      LETTERS[j],
      res$class_sizes[j],
      res$avg_post_prob[j],
      res$apcp_modal[j],
      res$occ[j]
    ))
  }
  
  cat(sprintf("  Relative entropy: %.4f\n", res$entropy))
}

# ================================================================= #
#### LR TEST 1: Do the membership covariables add value?         ####
# Comparing C=2 sin covar (model_2cl, k=12) vs
#         C=2 con covar (model_2cl_final, k=15)
# ================================================================= #

cat("\n", strrep("=", 70), "\n")
cat("  LR TESTS\n")
cat(strrep("=", 70), "\n")

# --- LR Test 1: membership covariates ---
ll_r1  <- as.numeric(model_2cl$LLout[1])        # restringded: without covar
ll_ur1 <- as.numeric(model_2cl_final$LLout[1])  # not restringed: with covar
k_r1   <- length(model_2cl$estimate)
k_ur1  <- length(model_2cl_final$estimate)
df1    <- k_ur1 - k_r1
LR1    <- -2 * (ll_r1 - ll_ur1)
pval1  <- pchisq(LR1, df = df1, lower.tail = FALSE)

cat("\nLR Test 1: Membership covariates (H0: all gamma = 0)\n")
cat(sprintf("  LL restringed    (sin covar, k=%d): %10.4f\n", k_r1,  ll_r1))
cat(sprintf("  LL not restringed (con covar, k=%d): %10.4f\n", k_ur1, ll_ur1))
cat(sprintf("  LR statistic: %.4f\n", LR1))
cat(sprintf("  df          : %d\n",   df1))
cat(sprintf("  p-value     : %.2e\n", pval1))

# --- LR Test 2: C=2 vs C=3 (auxiliary; p-value) ---
ll_r2  <- as.numeric(model_2cl_final$LLout[1])
ll_ur2 <- as.numeric(model_3cl$LLout[1])
k_r2   <- length(model_2cl_final$estimate)
k_ur2  <- length(model_3cl$estimate)
df2    <- k_ur2 - k_r2
LR2    <- -2 * (ll_r2 - ll_ur2)
pval2  <- pchisq(LR2, df = df2, lower.tail = FALSE)

cat("\nLR Test 2: Number of classes C=2 vs C=3\n")
cat("  [NOTE: mild p-value — parameter in the limit under H0]\n")
cat(sprintf("  LL C=2 (con covar, k=%d): %10.4f\n", k_r2,  ll_r2))
cat(sprintf("  LL C=3 (con covar, k=%d): %10.4f\n", k_ur2, ll_ur2))
cat(sprintf("  LR statistic: %.4f\n", LR2))
cat(sprintf("  df          : %d\n",   df2))
cat(sprintf("  p-valor chi2: %.2e  (conservador)\n", pval2))
cat(sprintf("  Convergence C=3: SINGULAR — SE not available \n"))
cat(sprintf("  Main criteria: BIC selects C=2 .\n"))

cat("\n WARNING:\n")
cat("  C=3: singular convergence — no valid covariance matrix.\n")
cat("  C=4: BHHH singular —  no standard error.\n")
cat("  Solo C=2 negative hessian and valid SE.\n")
cat("  Selected model: C=2.\n")

```

### Class selection graphics

```{r}
# ================================================================= #
#### GRÁFICOS                                                    ####
# ================================================================= #

theme_ac <- function() {
  theme_classic(base_size = 11, base_family = "serif") +
    theme(
      plot.title    = element_text(face = "bold", size = 12),
      plot.subtitle = element_text(size = 10, color = "grey40"),
      plot.caption  = element_text(size = 8,  color = "grey50"),
      legend.position    = "bottom",
      legend.title       = element_blank(),
      panel.grid.major.y = element_line(color = "grey92",
                                        linewidth = 0.4)
    )
}

# ---- Plot A: convergence criteria  ------------ #

p_fit <- ggplot(fit_table, aes(x = Classes)) +
  # Problematic convergence shadowed
  annotate("rect", xmin = 2.5, xmax = 4.5,
           ymin = -Inf, ymax = Inf,
           fill = "grey95", alpha = 0.7) +
  annotate("text", x = 3.5,
           y = max(fit_table$BIC) * 1.0005,
           label = "Singular convergence\n(no valid SE)",
           size = 2.8, hjust = 0.5, color = "grey55",
           family = "serif", fontface = "italic") +
  geom_line(aes(y = BIC, color = "BIC"), linewidth = 1.0) +
  geom_line(aes(y = AIC, color = "AIC"), linewidth = 1.0,
            linetype = "dashed") +
  # Puntos convergidos
  geom_point(data = filter(fit_table,  Converged),
             aes(y = BIC, color = "BIC"), size = 4) +
  geom_point(data = filter(fit_table,  Converged),
             aes(y = AIC, color = "AIC"), size = 4, shape = 17) +
  # Puntos NO convergidos (X)
  geom_point(data = filter(fit_table, !Converged),
             aes(y = BIC, color = "BIC"),
             size = 4, shape = 4, stroke = 1.8) +
  geom_point(data = filter(fit_table, !Converged),
             aes(y = AIC, color = "AIC"),
             size = 4, shape = 4, stroke = 1.8) +
  # vERTICAL LINE: SELECTED MODEL
  geom_vline(xintercept = 2, linetype = "dotted",
             color = "#2C5F8A", linewidth = 0.8) +
  annotate("text", x = 2.08,
           y = min(fit_table$AIC) +
             0.55 * (max(fit_table$BIC) - min(fit_table$AIC)),
           label = "Selected\n(C = 2)",
           size = 3, hjust = 0, color = "#2C5F8A",
           family = "serif", fontface = "bold") +
  scale_color_manual(values = c("BIC" = "#2C5F8A",
                                "AIC" = "#C0392B")) +
  scale_x_continuous(breaks = 2:4) +
  labs(
    title    = "A. Information criteria by number of classes",
    subtitle = paste0(
      "Filled/triangle = favorable convergence;  ",
      "\u2717 = singular convergence (no SE available)"),
    x        = "Number of latent classes (C)",
    y        = "Value",
    caption  = paste0(
      "Selected model: C = 2 \u2014 only specification with ",
      "negative definite Hessian and valid covariance matrix.")
  ) +
  theme_ac()

print(p_fit)
ggsave("LC_sel_A_criteria.pdf",  p_fit, width = 14, height = 9,
       units = "cm", device = cairo_pdf)
ggsave("LC_sel_A_criteria.tiff", p_fit, width = 14, height = 9,
       units = "cm", dpi = 600, compression = "lzw")
cat("Plot A saved\n")

# ---- Plot B: entropy ------------------------------------------- #

p_entropy <- ggplot(fit_table, aes(x = Classes, y = Entropy)) +
  geom_line(color = "#2C5F8A", linewidth = 1.0) +
  geom_point(data = filter(fit_table,  Converged),
             color = "#2C5F8A", size = 3.5) +
  geom_point(data = filter(fit_table, !Converged),
             color = "#2C5F8A", size = 3.5,
             shape = 4, stroke = 1.5) +
  geom_hline(yintercept = 0.6, linetype = "dashed",
             color = "grey50", linewidth = 0.6) +
  annotate("text", x = max(fit_table$Classes), y = 0.615,
           label = "Threshold = 0.60", size = 3,
           hjust = 1, color = "grey50", family = "serif") +
  scale_x_continuous(breaks = 2:4) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(
    title    = "B. Relative entropy by number of classes",
    subtitle = paste0(
      "All specifications exceed the 0.60 threshold;  ",
      "\u2717 = singular convergence"),
    x        = "Number of latent classes (C)",
    y        = "Relative entropy",
    caption  = "Dashed line: acceptable threshold (\u2265 0.60)."
  ) +
  theme_ac()

print(p_entropy)
ggsave("LC_sel_B_entropy.pdf",  p_entropy, width = 12, height = 9,
       units = "cm", device = cairo_pdf)
ggsave("LC_sel_B_entropy.tiff", p_entropy, width = 12, height = 9,
       units = "cm", dpi = 600, compression = "lzw")
cat("Plot B guardado.\n")

# ---- Plot C:class sizes  ----------------------------------- #

sizes_long <- bind_rows(lapply(results, function(res) {
  tibble(
    Classes = res$C,
    Class   = paste0("Class ", LETTERS[seq_len(res$C)]),
    Size    = res$class_sizes
  )
}))

p_sizes <- ggplot(sizes_long,
                  aes(x = factor(Classes), y = Size, fill = Class)) +
  geom_col(position = "stack", width = 0.5, alpha = 0.85) +
  geom_hline(yintercept = 0.05, linetype = "dashed",
             color = "grey40", linewidth = 0.5) +
  annotate("text", x = 0.6, y = 0.065, label = "5% min.",
           size = 3, hjust = 0, color = "grey40",
           family = "serif") +
  scale_y_continuous(labels = percent_format(),
                     limits = c(0, 1),
                     breaks = seq(0, 1, 0.2)) +
  scale_fill_brewer(palette = "Blues", direction = -1) +
  labs(
    title    = "Class size distribution",
    subtitle = "Unconditional class membership probabilities",
    x        = "Number of latent classes (C)",
    y        = "Class share",
    fill     = NULL,
    caption  = "Dashed line: 5% minimum class size guideline."
  ) +
  theme_ac()

print(p_sizes)
ggsave("LC_sel_C_sizes.pdf",  p_sizes, width = 12, height = 9,
       units = "cm", device = cairo_pdf)
ggsave("LC_sel_C_sizes.tiff", p_sizes, width = 12, height = 9,
       units = "cm", dpi = 600, compression = "lzw")
cat("Plot C saved\n")

# ---- Plot D:posterior probability and class sizes ----- #

post_long <- bind_rows(lapply(results, function(res) {
  tibble(
    Classes   = res$C,
    Class     = paste0("Class ", LETTERS[seq_len(res$C)]),
    PostProb  = res$avg_post_prob,
    ClassSize = res$class_sizes
  )
}))

p_post <- ggplot(post_long,
                 aes(x = factor(Classes), y = PostProb,
                     fill = Class)) +
  geom_col(position = position_dodge(width = 0.65),
           width = 0.5, alpha = 0.85) +
  # Hyphen in position of inconditional class size
  # If bar  ≈ hyohen → posteriors ≈ priors → low discrimination
  geom_point(aes(y = ClassSize, group = Class),
             position = position_dodge(width = 0.65),
             shape = 95, size = 7, color = "grey25") +
  geom_hline(yintercept = 0.7, linetype = "dashed",
             color = "grey40", linewidth = 0.5) +
  annotate("text", x = 0.6, y = 0.715, label = "0.70 min.",
           size = 3, hjust = 0, color = "grey40",
           family = "serif") +
  scale_y_continuous(limits = c(0, 1),
                     breaks = seq(0, 1, 0.2)) +
  scale_fill_brewer(palette = "Blues", direction = -1) +
  labs(
    title    = "D. Average posterior class probabilities",
    subtitle = paste0(
      "Bars = avg. posterior probability;  ",
      "\u2014 = unconditional class size"),
    x        = "Number of latent classes (C)",
    y        = "Average posterior probability",
    fill     = NULL,
    caption  = paste0(
      "When bars \u2248 dashes, posteriors \u2248 priors: ",
      "limited individual-level discrimination.  ",
      "Dashed line: minimum threshold (\u2265 0.70).")
  ) +
  theme_ac()

print(p_post)
ggsave("LC_sel_D_postprob.pdf",  p_post, width = 12, height = 9,
       units = "cm", device = cairo_pdf)
ggsave("LC_sel_D_postprob.tiff", p_post, width = 12, height = 9,
       units = "cm", dpi = 600, compression = "lzw")
cat("Plot D guardado.\n")

# ================================================================= #
#### TABLE LATEX FINAL                                           ####
# ================================================================= #

cat("\n\n% ===== TABLe LATEX FINAL =====\n")
cat("\\begin{table}[h!]\n\\centering\\small\n")
cat("\\caption{Latent class model selection: fit statistics and ",
    "classification diagnostics}\n")
cat("\\label{tab:LC_selection}\n")
cat("\\begin{threeparttable}\n")
cat("\\begin{tabular}{crrrrrrrc}\n\\toprule\n")
cat("$C$ & LL & $k$ & AIC & BIC & $\\Delta$BIC & ",
    "$\\rho^2_0$ & $\\rho^2_c$ & Convergence \\\\\n\\midrule\n")

for (i in seq_len(nrow(fit_table))) {
  r         <- fit_table[i, ]
  conv_rows <- fit_table[fit_table$Converged, ]
  is_best   <- r$Converged && r$BIC == min(conv_rows$BIC)
  marker    <- if (is_best) "$^{\\dagger}$" else ""
  conv_str  <- if (r$Converged) "Favorable"
               else "Singular$^{a}$"
  cat(sprintf(
    "%d & %.2f & %d & %.2f & %.2f%s & %.2f & %.4f & %.4f & %s \\\\\n",
    r$Classes, r$LL, r$npar,
    r$AIC, r$BIC, marker,
    r$delta_BIC, r$Rho2_0, r$Rho2_c,
    conv_str))
}

cat("\\bottomrule\n\\end{tabular}\n")
cat("\\begin{tablenotes}[flushleft]\\footnotesize\n")
cat("\\item \\textit{Notes:} $C$ = number of latent classes; ")
cat("$k$ = estimated parameters; LL = log-likelihood at convergence; ")
cat("$\\Delta$BIC = difference relative to minimum BIC; ")
cat("$\\rho^2_0$ computed against the equal-probability null model ")
cat("($\\text{LL}_0 = -N_{obs} \\log 2$); ")
cat("Entropy = relative entropy (higher = better class separation). ")
cat("$^{\\dagger}$ Selected specification: lowest BIC among ")
cat("specifications with favorable convergence and valid covariance matrix. ")
cat("$^{a}$ Singular or degenerate convergence: parameter estimates ")
cat("are available but standard errors cannot be computed.\n")
cat("\\end{tablenotes}\n")
cat("\\end{threeparttable}\n\\end{table}\n")
cat(sprintf(
  "\n%% LR Test (membership covariates): chi2(%d) = %.4f, p = %.2e\n",
  df1, LR1, pval1))
cat(sprintf(
  "%% LR Test (C=2 vs C=3, auxiliar):  chi2(%d) = %.4f, p = %.2e [conservador; C=3 singular]\n",
  df2, LR2, pval2))

cat("\nProcess finished.\n")
```




### Final 2 classes model


```{r}
rm(list = ls())
data_path <- file.path("data", "Analisis.xlsx")


apollo_initialise()

apollo_control = list(
  modelName       = "LC_2cl_TPB_Surv_final",
  modelDescr      = "LC 2 classes with TPB + Registry, restringed price",
  indivID         = "Id",
  mixing          = FALSE,
  nCores          = 4,
  outputDirectory = "output"
)

database <- as.data.frame(
  read_excel(data_path,
             sheet = "Todo_II") %>%
    mutate(
      Project    = ifelse(Optout == 1 & Project  == -4, 0, Project),
      Specie     = ifelse(Optout == 1 & Specie   == -4, 0, Specie),
      Location   = ifelse(Optout == 1 & Location == -4, 0, Location),
      Price      = ifelse(Optout == 1 & Price    == -4, 0, Price),
      ASC_accept = as.numeric(ifelse(Optout == 1, 0, 1)),
      alt_id     = ifelse(Optout == 0, 1, 2),
      chid       = paste(Id, ParentQuestionId, QuestionId, sep = "_")
    ) %>%
    group_by(chid) %>%
    mutate(choiceVar = alt_id[which(Choice == 1)][1]) %>%
    ungroup()
)

# ----------------------------------------------------------------- #
# Price - Guaranteed negative sign
# ----------------------------------------------------------------- #
apollo_beta = c(
  asc_a          =  1.0,
  asc_b          = -1.0,

  b_Project_a    =  0.10,
  b_Specie_a     =  0.20,
  b_Location_a   =  0.30,
  b_Price_log_a  = -3.91,   # -exp(-3.91) ≈ -0.020

  b_Project_b    = -0.05,
  b_Specie_b     =  0.10,
  b_Location_b   =  0.10,
  b_Price_log_b  = -3.91,

  delta_a        =  0.0,
  delta_b        =  0.0,    # referenec, fixed

  gamma_Att_a    =  0.0,
  gamma_PBC_a    =  0.0,
  gamma_SN_a     =  0.0,
  gamma_Surv_a   =  0.0
  #gamma_Pyme_a   =  0.0
)

# Class B = reference; only gamma for Class A
apollo_fixed = c("delta_b")

# ----------------------------------------------------------------- #
apollo_lcPars = function(apollo_beta, apollo_inputs) {
  lcpars = list()

  lcpars[["asc"]]        = list(asc_a,        asc_b)
  lcpars[["b_Project"]]  = list(b_Project_a,  b_Project_b)
  lcpars[["b_Specie"]]   = list(b_Specie_a,   b_Specie_b)
  lcpars[["b_Location"]] = list(b_Location_a, b_Location_b)
  lcpars[["b_Price"]]    = list(-exp(b_Price_log_a),
                                 -exp(b_Price_log_b))

  V_class = list()
  V_class[["class_a"]] = delta_a +
    gamma_Att_a  * Attitudes       +
    gamma_PBC_a  * PerceivedBehavC +
    gamma_SN_a   * SubjectiveNorms +
    gamma_Surv_a * Survey          #+
    #gamma_SME_a * SME

  V_class[["class_b"]] = delta_b   # = 0, reference

  lcpars[["pi_values"]] = apollo_classAlloc(list(
    classes       = c(class_a = 1, class_b = 2),
    utilities     = V_class,
    componentName = "classAlloc"
  ))

  return(lcpars)
}

# ----------------------------------------------------------------- #
apollo_inputs = apollo_validateInputs()

apollo_probabilities = function(apollo_beta, apollo_inputs,
                                functionality = "estimate") {
  apollo_attach(apollo_beta, apollo_inputs)
  on.exit(apollo_detach(apollo_beta, apollo_inputs))

  P = list()
  mnl_settings = list(
    alternatives = c(project = 1, optout = 2),
    avail        = list(project = 1, optout = 1),
    choiceVar    = choiceVar
  )

  for (s in 1:2) {
    V = list()
    V[["project"]] = asc[[s]]        * ASC_accept +
                     b_Project[[s]]  * Project    +
                     b_Specie[[s]]   * Specie     +
                     b_Location[[s]] * Location   +
                     b_Price[[s]]    * Price
    V[["optout"]]  = 0

    mnl_settings$utilities     = V
    mnl_settings$componentName = paste0("Class_", s)

    P[[paste0("Class_", s)]] = apollo_mnl(mnl_settings, functionality)
    P[[paste0("Class_", s)]] = apollo_panelProd(
      P[[paste0("Class_", s)]], apollo_inputs, functionality)
  }

  lc_settings  = list(inClassProb = P, classProb = pi_values)
  P[["model"]] = apollo_lc(lc_settings, apollo_inputs, functionality)
  P = apollo_prepareProb(P, apollo_inputs, functionality)
  return(P)
}

# ----------------------------------------------------------------- #
model_2cl_final = apollo_estimate(
  apollo_beta, apollo_fixed,
  apollo_probabilities, apollo_inputs,
  estimate_settings = list(
    estimationRoutine = "bgw",
    maxIterations     = 2000,
    writeIter         = FALSE,
    numDeriv          = TRUE    
  )
)

apollo_modelOutput(model_2cl_final,
                   modelOutput_settings = list(printPVal = TRUE))
apollo_saveOutput(model_2cl_final)
apollo_sink()

# ----------------------------------------------------------------- #
# DIAGNOSTICS
# ----------------------------------------------------------------- #
cat("\nNAs en robvarcov:", sum(is.na(model_2cl_final$robvarcov)), "\n")

unconditionals = apollo_lcUnconditionals(model_2cl_final,
                                          apollo_probabilities,
                                          apollo_inputs)
n_cl        = length(unconditionals[["pi_values"]])
class_sizes = colMeans(matrix(unlist(unconditionals[["pi_values"]]),
                               ncol = n_cl, byrow = FALSE))
cat("\n=== Classes sizes ===\n")
print(round(class_sizes, 4))

conditionals = apollo_lcConditionals(model_2cl_final,
                                      apollo_probabilities,
                                      apollo_inputs)
class_modal = apply(conditionals[, 2:3], 1, which.max)
cat("\n=== modal asignation ===\n")
print(table(class_modal))

for (var in c("Segment","Survey","SME","SustDepart",
              "Attitudes","PerceivedBehavC","SubjectiveNorms","Risk")) {
  if (var %in% names(database)) {
    vn = apollo_firstRow(database[[var]], apollo_inputs)
    pv = colSums(vn * conditionals[, 2:3]) / colSums(conditionals[, 2:3])
    cat(sprintf("\n%-20s | A: %6.3f | B: %6.3f\n", var, pv[1], pv[2]))
  }
}

# ----------------------------------------------------------------- #
# WTP
# ----------------------------------------------------------------- #
b = model_2cl_final$estimate

for (cl in c("a", "b")) {
  price_real = -exp(b[paste0("b_Price_log_", cl)])
  cat(sprintf("\n=== WTP Clase %s (precio = %.5f) ===\n",
              toupper(cl), price_real))
  cat("  ASC:     ", round(-b[paste0("asc_",       cl)] / price_real, 2), "\n")
  cat("  Project: ", round(-b[paste0("b_Project_", cl)] / price_real, 2), "\n")
  cat("  Specie:  ", round(-b[paste0("b_Specie_",  cl)] / price_real, 2), "\n")
  cat("  Location:", round(-b[paste0("b_Location_",cl)] / price_real, 2), "\n")
}

# ----------------------------------------------------------------- #
# KRINSKY-ROBB
# ----------------------------------------------------------------- #
if (sum(is.na(model_2cl_final$robvarcov)) == 0) {

  set.seed(2025)
  S      = 10000
  V_vcov = model_2cl_final$robvarcov
  b_kr   = model_2cl_final$estimate[rownames(V_vcov)]
  draws  = MASS::mvrnorm(n = S, mu = b_kr, Sigma = V_vcov)

  price_a = -exp(draws[, "b_Price_log_a"])
  price_b = -exp(draws[, "b_Price_log_b"])

  wtp_list = list(
    ASC_a      = -draws[, "asc_a"]        / price_a,
    Project_a  = -draws[, "b_Project_a"]  / price_a,
    Specie_a   = -draws[, "b_Specie_a"]   / price_a,
    Location_a = -draws[, "b_Location_a"] / price_a,
    ASC_b      = -draws[, "asc_b"]        / price_b,
    Project_b  = -draws[, "b_Project_b"]  / price_b,
    Specie_b   = -draws[, "b_Specie_b"]   / price_b,
    Location_b = -draws[, "b_Location_b"] / price_b
  )

  cat("\n=== Krinsky-Robb WTP [mean, median, q025, q975] ===\n")
  kr_results = lapply(wtp_list, function(x) {
    c(mean   = round(mean(x,   na.rm = TRUE), 2),
      median = round(median(x, na.rm = TRUE), 2),
      q025   = round(quantile(x, 0.025, na.rm = TRUE), 2),
      q975   = round(quantile(x, 0.975, na.rm = TRUE), 2))
  })
  print(do.call(rbind, kr_results))

} else {
  cat("\nATENCIÓN: robvarcov has NAs after numDeriv=TRUE.\n")
  cat(" Model may have structural identification problems.\n")
}

# Information criteria
ll   = model_2cl_Pyme$LLout[1]
  k    = length(model_2cl_Pyme$estimate)
n_cs = length(unique(database$chid))
cat("\n=== Criterios de información ===\n")
cat("LL  :", round(ll,  2), "\n")
cat("AIC :", round(-2*ll + 2*k,         2), "\n")
cat("BIC :", round(-2*ll + k*log(n_cs), 2), "\n")
cat("npar:", k, "\n")

# Information criteria
ll   = model_2cl_final$LLout[1]
  k    = length(model_2cl_final$estimate)
n_cs = length(unique(database$chid))
cat("\n=== Criterios de información ===\n")
cat("LL  :", round(ll,  2), "\n")
cat("AIC :", round(-2*ll + 2*k,         2), "\n")
cat("BIC :", round(-2*ll + k*log(n_cs), 2), "\n")
cat("npar:", k, "\n")

apollo_sink()
```



### Data visualisation

```{r}
col_A   <- "#79CDCD"
col_B   <- "#6E8B3D"
col_mid <- "#7F8C8D"

theme_academic <- function() {
  theme_classic(base_size = 11, base_family = "serif") +
    theme(
      plot.title    = element_text(face = "bold", size = 12, hjust = 0),
      plot.subtitle = element_text(size = 10, color = "grey40", hjust = 0),
      axis.title    = element_text(size = 10),
      axis.text     = element_text(size = 9),
      legend.position  = "bottom",
      legend.title     = element_blank(),
      legend.text      = element_text(size = 9),
      panel.grid.major.y = element_line(color = "grey92", linewidth = 0.4),
      strip.text       = element_text(face = "bold", size = 10),
      plot.caption     = element_text(size = 8, color = "grey50", hjust = 0)
    )
}

# ----------------------------------------------------------------- #
# DATA
# ----------------------------------------------------------------- #

wtp_data <- tibble(
  class     = rep(c("Class A\n(ESG-oriented)",
                    "Class B\n(Capacity constrained)"), each = 3),
  attribute = rep(c("Project type", "Tree species", "Location"), 2),
  mean      = c(3.9,   5.9,  15.1,   -4.6,  5.9,   2.4),
  q025      = c(-0.5,  2.0,   9.9,  -13.7, -1.8,  -5.9),
  q975      = c( 8.2, 10.3,  20.8,    1.6, 20.6,  15.6)
) %>%
  mutate(
    attribute = factor(attribute,
                       levels = c("Project type", "Tree species", "Location")),
    sig = case_when(
      q025 > 0 | q975 < 0 ~ TRUE,
      TRUE                 ~ FALSE
    )
  )

membership_data <- tibble(
  parameter = c("Registry\nmembership", "PBC", "Attitudes",
                "Subjective\nNorms", "Intercept"),
  estimate  = c(1.946,  0.341,  0.216,  0.207, -1.298),
  se_rob    = c(0.253,  0.121,  0.120,  0.127,  0.221),
  sig       = c("***",  "***",  "**",   "*",    "***")
) %>%
  mutate(
    lo = estimate - 1.96 * se_rob,
    hi = estimate + 1.96 * se_rob,
    parameter = factor(parameter,
                       levels = rev(c("Registry\nmembership", "PBC",
                                      "Attitudes", "Subjective\nNorms",
                                      "Intercept")))
  )

posterior_data <- tibble(
  variable = rep(c("Awareness", "Registry\nmembership",
                   "Sustainability\ndept.", "SME",
                   "Service sector",
                   "Attitudes\n(mean)", "PBC\n(mean)",
                   "Subj. Norms\n(mean)"), 2),
  class    = rep(c("Class A\n(ESG-oriented)",
                   "Class B\n(Price-oriented)"), each = 8),
  value    = c(
    # Class A
    0.706, 0.661, 0.462, 0.588, NA,
    0.181, 0.219, 0.189,

    # Class B
    0.870, 0.240, 0.209, 0.802, NA,
    -0.121, -0.146, -0.126
  ),
  type     = rep(c("binary", "binary", "binary", "binary", "binary",
                   "continuous", "continuous", "continuous"), 2)
)

size_data <- tibble(
  class   = c("Class A\n(ESG-oriented)", "Class B\n(Price-oriented)"),
  share   = c(0.3997, 0.6003),
  n_modal = c(203, 319),
  label   = c("40.0%\n(N = 203)", "60.0%\n(N = 319)")
) %>%
  mutate(
    ymax      = cumsum(share),
    ymin      = lag(ymax, default = 0),
    label_pos = (ymax + ymin) / 2
  )

# ----------------------------------------------------------------- #
# PLOT A — WTP by class at IC95%
# ----------------------------------------------------------------- #

p1 <- ggplot(wtp_data,
             aes(x = attribute, y = mean,
                 color = class, shape = sig)) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "grey60", linewidth = 0.5) +
  geom_errorbar(aes(ymin = q025, ymax = q975),
                width = 0.15, linewidth = 0.7,
                position = position_dodge(width = 0.5)) +
  geom_point(size = 3.5, fill = "white", stroke = 1.2,
             position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c(col_A, col_B)) +
  scale_shape_manual(values = c(`FALSE` = 21, `TRUE` = 19),
                     guide  = "none") +
  scale_y_continuous(labels = label_dollar(prefix = "€", suffix = ""),
                     breaks = seq(-20, 25, by = 5)) +
  labs(
    title    = "A. Willingness to Pay by preference class",
    subtitle = "Mean WTP with 95% Krinsky–Robb confidence intervals (€ per tCO₂)",
    x        = NULL,
    y        = "WTP (€ per tCO₂)",
    caption  = "Filled points indicate statistically significant WTP (CI excludes zero)."
  ) +
  theme_academic()

print(p1)

ggsave("LC_plot_A_WTP.pdf",  plot = p1,
       width = 14, height = 9, units = "cm", device = cairo_pdf)
ggsave("LC_plot_A_WTP.tiff", plot = p1,
       width = 14, height = 9, units = "cm", dpi = 600,
       compression = "lzw")

cat("Plot A saved\n")

# ----------------------------------------------------------------- #
# PLOT B — Concomitant model forest plot
# ----------------------------------------------------------------- #

p2 <- ggplot(membership_data,
             aes(x = estimate, y = parameter)) +
  geom_vline(xintercept = 0, linetype = "dashed",
             color = "grey60", linewidth = 0.5) +
  geom_errorbarh(aes(xmin = lo, xmax = hi),
                 height = 0.25, linewidth = 0.7,
                 color = col_A) +
  geom_point(size = 3.5, color = col_A) +
  geom_text(aes(x = hi + 0.12, label = sig),
            size = 3.5, hjust = 0, color = "grey30") +
  scale_x_continuous(limits = c(-2.2, 3.0),
                     breaks = seq(-2, 3, by = 1)) +
  labs(
    title    = "A. Class-membership model",
    subtitle = "Log-odds of Class A (ESG-oriented) vs. Class B, with 95% CI",
    x        = "Coefficient estimate",
    y        = NULL,
    caption  = "* p < 0.10, ** p < 0.05, *** p < 0.01 (robust standard errors)."
  ) +
  theme_academic()

print(p2)

ggsave("LC_plot_A_membership.pdf",  plot = p2,
       width = 12, height = 9, units = "cm", device = cairo_pdf)
ggsave("LC_plot_B_membership.tiff", plot = p2,
       width = 12, height = 9, units = "cm", dpi = 600,
       compression = "lzw")

cat("Plot B saved\n")

# ----------------------------------------------------------------- #
# PLOT C — posterior profile (binary variables + TPB)
# ----------------------------------------------------------------- #



post_binary <- posterior_data %>%
  filter(type == "binary", !is.na(value)) %>%
  mutate(variable = factor(variable,
                           levels = c("Awareness",
                                      "Registry\nmembership",
                                      "Sustainability\ndept.",
                                      "SME",
                                      "Service sector")))
post_cont <- posterior_data %>%
  filter(type == "continuous") %>%
  mutate(variable = factor(variable,
                           levels = c("Attitudes\n(mean)",
                                      "PBC\n(mean)",
                                      "Subj. Norms\n(mean)")))

p3a <- ggplot(post_binary,
              aes(x = variable, y = value, fill = class)) +
  geom_col(position = position_dodge(width = 0.6),
           width = 0.5, alpha = 0.85) +
  scale_fill_manual(values = c(col_A, col_B)) +
  scale_y_continuous(labels = percent_format(),
                     limits = c(0, 1),
                     breaks = seq(0, 1, by = 0.2)) +
  labs(title    = " ",
       subtitle = "Firm characteristics (share per class)",
       x = NULL, y = "Share") +
  theme_academic() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 25, hjust = 1)
  )

p3b <- ggplot(post_cont,
              aes(x = variable, y = value, fill = class)) +
  geom_col(position = position_dodge(width = 0.6),
           width = 0.5, alpha = 0.85) +
  geom_hline(yintercept = 0, linewidth = 0.4, color = "grey40") +
  scale_fill_manual(values = c(col_A, col_B)) +
  scale_y_continuous(limits = c(-0.25, 0.45),
                     breaks = seq(-0.2, 0.4, by = 0.1)) +
  labs(subtitle = "TPB construct means (standardised scores)",
       x = NULL, y = "Mean score (z)") +
  theme_academic() +
  theme(legend.position  = "none",
        plot.subtitle    = element_text(size = 10, color = "grey40"))

p3 <- p3a / p3b +
  plot_layout(heights = c(1, 0.8), guides = "collect") &
  theme(legend.position = "bottom")

print(p3)

ggsave("LC_plot_C_profile.pdf",  plot = p3,
       width = 16, height = 14, units = "cm", device = cairo_pdf)

ggsave("LC_plot_C_profile.tiff", plot = p3,
       width = 16, height = 14, units = "cm", dpi = 600,
       compression = "lzw")

cat("Plot C guardado.\n")

# ----------------------------------------------------------------- #
# PLOT D — class size
# ----------------------------------------------------------------- #

p4 <- ggplot(size_data,
             aes(ymax = ymax, ymin = ymin,
                 xmax = 4, xmin = 2.5, fill = class)) +
  geom_rect(alpha = 0.9) +
  geom_text(aes(x = 3.25, y = label_pos, label = label),
            size = 3.5, fontface = "bold", color = "white",
            family = "serif") +
  scale_fill_manual(values = c(col_A, col_B)) +
  coord_polar(theta = "y") +
  xlim(c(1, 4)) +
  labs(title    = "D. Class size distribution",
       subtitle = "Unconditional class membership probabilities") +
  theme_void(base_family = "serif") +
  theme(
    plot.title    = element_text(face = "bold", size = 12,
                                 hjust = 0.5,
                                 margin = margin(b = 4)),
    plot.subtitle = element_text(size = 10, color = "grey40",
                                 hjust = 0.5),
    legend.position = "none"
  )

print(p4)

ggsave("LC_plot_D_classsize.pdf",  plot = p4,
       width = 10, height = 10, units = "cm", device = cairo_pdf)
ggsave("LC_plot_D_classsize.tiff", plot = p4,
       width = 10, height = 10, units = "cm", dpi = 600,
       compression = "lzw")

cat("Plot D saved\n")
cat("\nTodos Graphics have been printed and saved.\n")
```
